---
title: "CD14 Data-set Anaysis with cleaning"
---

```{r}
library(ggplot2)
library(tibble)
library(zeallot)
library(COTAN)
library(Azimuth)
options(parallelly.fork.enable = TRUE)

outDir <- file.path(".", "CD14_NewSplit")
if (!dir.exists(outDir)) {
  dir.create(outDir)
}

inDir <- "."

setLoggingLevel(2)
setLoggingFile(file.path(outDir, "cd14_analysis.log"))
```

```{r}
library(Seurat)
dataset <- Read10X(file.path(inDir, "hg19"))
```

```{r}
sampleCondition <- "CD14_Monocytes"
cd14Obj <- COTAN(raw = dataset)
cd14Obj <- initializeMetaDataset(cd14Obj,
                                 GEO = "10X CD14+ Monocytes",
                                 sequencingMethod = "10X",
                                 sampleCondition = sampleCondition)

sampleCondition
```

Inspect cells' sizes

```{r}
cellSizePlot(cd14Obj)
```



Inspect the number of expressed genes per cell

```{r}
genesSizePlot(cd14Obj)
```



Check number of mitochondrial genes expressed in each cell

```{r}
mitGenesPattern <- "^[Mm][Tt]-"
getGenes(cd14Obj)[grep(mitGenesPattern, getGenes(cd14Obj))]

c(mitPlot, mitSizes) %<-%
  mitochondrialPercentagePlot(cd14Obj, genePrefix = mitGenesPattern)

plot(mitPlot)
```

Cells with a too high percentage of mitochondrial genes are likely dead (or at the last problematic) cells. So we drop them!

```{r}
mitPercThr <- 6
cd14Obj <- addElementToMetaDataset(cd14Obj, "Mitoc. perc. threshold", mitPercThr)

cells_to_rem <- rownames(mitSizes)[mitSizes[["mit.percentage"]] > mitPercThr]

cd14Obj <- dropGenesCells(cd14Obj, cells = cells_to_rem)

c(mitPlot, mitSizes) %<-%
  mitochondrialPercentagePlot(cd14Obj, genePrefix = mitGenesPattern)

plot(mitPlot)
```

Check the number of ribosomal genes expressed in each cell

```{r}
ribGenesPattern <- "^RP[SL]\\d+"
getGenes(cd14Obj)[grep(ribGenesPattern, getGenes(cd14Obj))]

c(ribPlot, ribSizes) %<-%
  mitochondrialPercentagePlot(cd14Obj, genePrefix = mitGenesPattern)

plot(ribPlot)
```

```{r}
scatterPlot(cd14Obj)
```


```{r}
cd14Obj <- clean(cd14Obj)

c(pcaCellsPlot, pcaCellsData, genesPlot, UDEPlot, nuPlot, zoomedNuPlot) %<-% cleanPlots(cd14Obj)

plot(pcaCellsPlot)
plot(genesPlot)

```


Visualize if all is ok:

```{r}

plot(UDEPlot)

plot(nuPlot)
plot(zoomedNuPlot)
```


Calculate genes' COEX

```{r, eval=FALSE, echo=TRUE}
Sys.time()

cd14Obj <- proceedToCoex(cd14Obj, calcCoex = T, cores = 5,
                         saveObj = TRUE, outDir = outDir)

```

```{r, eval=TRUE}
cd14Obj <- readRDS(file = file.path(outDir, paste0(sampleCondition, ".cotan.RDS")))
```

```{r}
gdiData <- calculateGDI(cd14Obj)
cd14Obj <- storeGDI(cd14Obj,genesGDI = gdiData)

genesToLabel <- head(rownames(gdiData[order(gdiData[["GDI"]],
                                            decreasing = TRUE), ]), n = 50L)

sort(genesToLabel)

gdiData[genesToLabel[[50L]], "GDI"]

gdiPlot <- GDIPlot(cd14Obj, GDIIn = gdiData, GDIThreshold = 1.4,
                   genes = list("Top 10 GDI genes" = genesToLabel[1L:10L]))

plot(gdiPlot)

Sys.time()
```

Save the COTAN object

```{r, eval=FALSE, echo=TRUE}
saveRDS(cd14Obj, file = file.path(outDir, paste0(sampleCondition, ".cotan.RDS")))
```

```{r}
cd14Obj <- readRDS(file = file.path(outDir, paste0(sampleCondition, ".cotan.RDS")))
```

```{r, eval=FALSE, echo=TRUE}
advChecker <- new("AdvancedGDIUniformityCheck")

c(splitClusters, splitCoexDF) %<-%
  cellsUniformClustering(objCOTAN = cd14Obj,
                         checker = advChecker,
                         cores = 5L,
                         genesSel = "HGDI",
                         saveObj = TRUE,
                         outDir = outDir)

cd14Obj <- addClusterization(cd14Obj, clName = "split",
                             clusters = splitClusters,
                             coexDF = splitCoexDF, override = FALSE)

table(splitClusters)
```
```{r}
cellSizePlot(cd14Obj, conditions = getClusters(cd14Obj, clName = "split"))
```

```{r}
clustersSummaryPlot(cd14Obj, clName = "split")
```

```{r}
cellsUMAPPlot(cd14Obj, clName = "split",
              genesSel = "HGDI", dataMethod = "AdjBinarized")
```

```{r, echo=FALSE, eval=FALSE}
saveRDS(cd14Obj, file = file.path(outDir, paste0(sampleCondition, ".cotan.RDS")))
```

```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(objCOTAN = cd14Obj,
                            checkers = advChecker,
                            clusters = getClusters(cd14Obj, clName = "split"),
                            cores = 5,
                            batchSize = 10,
                            saveObj = TRUE,
                            outDir =  outDir)

cd14Obj <- addClusterization(cd14Obj, clName = "merge",
                             clusters = mergedClusters,
                             coexDF = mergedCoexDF,
                             override = FALSE)

table(mergedClusters)

table(splitClusters, mergedClusters)
```


```{r, eval=TRUE}
mergedClusters <- getClusters(cd14Obj, clName = "merge")

table(mergedClusters)

table(splitClusters, mergedClusters)
```

```{r}
cellSizePlot(cd14Obj, conditions = getClusters(cd14Obj, clName = "merge"))
```

```{r}
scatterPlot(cd14Obj, conditions = getClusters(cd14Obj, clName = "merge"))
```

```{r}
clustersSummaryPlot(cd14Obj, clName = "merge")
```

```{r}
cellsUMAPPlot(cd14Obj, clName = "merge",
              genesSel = "HGDI", dataMethod = "AdjBinarized")
```


```{r, eval=FALSE}
saveRDS(cd14Obj, file = file.path(outDir, paste0(sampleCondition, ".cotan.RDS")))
```

```{r}
Sys.time()
```


```{r}
sessionInfo()
```
