---
title: "FDR analysis - Results - thresholds true 3% 10% "
author: "Silvia Giulia Galfrè"
date: "`r Sys.Date()`"
output: html_document
---

## Preamble

```{r}
library(COTAN)
library(pROC)
options(parallelly.fork.enable = TRUE)
library(Seurat)
library(monocle3)
library(reticulate)
library(cowplot)
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)

methods.color <- c("COTAN"="#F73604","Seurat"="#ABD9E9","Seurat_scTr"="#74ADD1","Seurat_Bimod"="#4575B4", "Monocle"="#DAABE9", "Scanpy"="#7F9B5C" )

dirOut <- "Results/FDR/"
if (!dir.exists(dirOut)) {
  dir.create(dirOut)
}

dataSetDir <- "Data/MouseCortexFromLoom/FDR/MergedClusters_For_FDR/"
```

## Dataset composition

```{r}
datasets_csv <- read.csv(file.path(dataSetDir,"Cells_Usage_DataFrame.csv"),
                         row.names = 1
                        ) 

datasets_csv[1:3,]
```

### Define which genes are expressed

For each data set we need to define, independently from the DEA methods, which genes are specific for each cluster. So we need to define first which genes are expressed and which are not expressed. To do so we can take advantage from the fact that we have the original clusters from which the cells were sampled to create the artificial datasets. So looking to the original cluster we define as expressed all genes present in at least the 10% of cells and we define as not expressed the genes completely absent or expressed in less than 3% of cells.

Since these two thresholds can have a big influence on the tools performances we will test also others in other pages.

```{r}
file.presence <- readRDS("Data/MouseCortexFromLoom/FDR/Results/GenePresence_PerCluster.RDS")

for (file in list.files("Data/MouseCortexFromLoom/SingleClusterRawData/")) {
#  print(file)
  Code <- str_split(file,pattern = "_",simplify = T)[1]
  Time <- str_split(Code,pattern = "e",simplify = T)[2]
  Cluster <- str_split(Code,pattern = "e",simplify = T)[1]
  Cluster <- str_remove(Cluster,pattern = "Cl")
  Cluster <- paste0("E",Time,"-",Cluster)
  file.presence[,Cluster] <- "Absent"
  dataset.cl <- readRDS(file.path("Data/MouseCortexFromLoom/SingleClusterRawData/",
                                         file))
  number.cell.expressing <- rowSums(dataset.cl > 0)
  AbsentThreshold <- round(0.03*dim(dataset.cl)[2],digits = 0)
  PresenceThreshold <- round(0.1*dim(dataset.cl)[2],digits = 0)
  file.presence[names(number.cell.expressing[number.cell.expressing > AbsentThreshold]),Cluster] <- "Uncertain" 
  
  file.presence[names(number.cell.expressing[number.cell.expressing >= PresenceThreshold]),Cluster] <- "Present"
  print(Cluster)
  print(table(file.presence[,Cluster]))
  
  }
```

# 2 Clusters even

## 2_Clusters_even_near

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_near",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot,20)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan1, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat1, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle1, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy1, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
TwoClusters_even_near <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

TwoClusters_even_nearPL <- TwoClusters_even_near + 
  xlab("FPR") + 
  ylab("TPR")
TwoClusters_even_nearPL
```

## 2_Clusters_even_medium

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_medium",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan2, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat2, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle2, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy2, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Two_Clusters_even_medium <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Two_Clusters_even_mediumPL <- Two_Clusters_even_medium + xlab("FPR") + ylab("TPR")
Two_Clusters_even_mediumPL
```

## 2_Clusters_even_far

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_far",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan3, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat3, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle3, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy3, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Two_Clusters_even_far <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Two_Clusters_even_farPL <- Two_Clusters_even_far + xlab("FPR") + ylab("TPR")
Two_Clusters_even_farPL
```

# 2 clusters uneven

## 2_Clusters_uneven_near

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_near",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan4, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat4, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle4, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy4, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Two_Clusters_uneven_near <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Two_Clusters_uneven_nearPL <- Two_Clusters_uneven_near + xlab("FPR") + ylab("TPR")
Two_Clusters_uneven_nearPL
```

## 2_Clusters_uneven_medium

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_medium",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan5, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat5, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle5, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy5, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Two_Clusters_uneven_medium <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Two_Clusters_uneven_mediumPL <- Two_Clusters_uneven_medium + xlab("FPR") + ylab("TPR")

Two_Clusters_uneven_mediumPL
```

## 2_Clusters_uneven_far

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_far",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan6, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat6, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle6, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy6, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Two_Clusters_uneven_far <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Two_Clusters_uneven_farPL <- Two_Clusters_uneven_far + xlab("FPR") + ylab("TPR")

lg <- get_legend(Two_Clusters_uneven_farPL)

Two_Clusters_uneven_farPL
```

# 3 clusters

## 3_Clusters_even

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_even",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan7, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat7, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle7, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy7, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Three_Clusters_even <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Three_Clusters_evenPL <- Three_Clusters_even + 
  xlab("FPR") + ylab("TPR")+theme(legend.position="none")
Three_Clusters_even
```

## 3_Clusters_uneven

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_uneven",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan8, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat8, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle8, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy8, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Three_Clusters_uneven <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

Three_Clusters_unevenPL <- Three_Clusters_uneven + xlab("FPR") + ylab("TPR")+theme(legend.position="none")

Three_Clusters_uneven
```

# 5 clusters

## 5_Clusters_uneven

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "5_Clusters_uneven",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.6
  
  file.presence.subset <- ground_truth
  
  for (col in 1:ncol(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset[,-col,drop = FALSE]) < 0.35  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]



head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes)) 
sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r cotan9, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  genesLambda <- getLambda(dataset)
  genesLambda <- as.data.frame(genesLambda)
  genesLambda$genes <- rownames(genesLambda)
  
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  
  onlyPositive.pVal.Cotan <- merge(onlyPositive.pVal.Cotan,genesLambda,by="genes")
  
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)


  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

cotan.threshold.pval <- quantile(onlyPositive.pVal.Cotan_tot[onlyPositive.pVal.Cotan_tot$value == 0,]$p_values,probs = 0.05)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat9, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

deaSeurat_tot_base <- deaSeurat_tot

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)


seurat.threshold.pval <- quantile(deaSeurat_tot_base[deaSeurat_tot_base$value == 0,]$p_val,probs = 0.05)
# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

deaSeurat_tot_scTr <- deaSeurat_tot

Seurat_scTr.threshold.pval <- quantile(deaSeurat_tot_scTr[deaSeurat_tot_scTr$value == 0,]$p_val,probs = 0.05)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Seurat Bimod

```{r, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

deaSeurat_tot_Bimod <- deaSeurat_tot

seurat_bimod.threshold.pval <- quantile(deaSeurat_tot_Bimod[deaSeurat_tot_Bimod$value == 0,]$p_val,probs = 0.05)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)
```

### ROC for Monocle

```{r monocle9, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

monocle.threshold.pval <- quantile(deaMonocle_tot[deaMonocle_tot$value == 0,]$marker_test_p_value,probs = 0.05)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from Scanpy

```{r Scanpy9, echo=TRUE, eval=TRUE}
deaScanpy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScanpy <- read.csv(file.path(dirOut,paste0(file.code,"Scanpy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScanpy[deaScanpy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScanpy$data_set <- subset.datasets_csv[ind,1]
  deaScanpy$set_number <- ind 
  deaScanpy_tot <- rbind(deaScanpy_tot, deaScanpy)
  
  
  }
 deaScanpy_tot <-  deaScanpy_tot[2:nrow(deaScanpy_tot),]
 
 deaScanpy_tot <- merge.data.frame(deaScanpy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScanpy_tot$value <- as.numeric(deaScanpy_tot$value)

scanpy.threshold.pval <- quantile(deaScanpy_tot[deaScanpy_tot$value == 0,]$pval,probs = 0.05)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScanpy <- roc(deaScanpy_tot$value, 1 - deaScanpy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScanpy)
```

### Summary ROC for all methods

```{r}
Five_Clusters <- ggroc(list(COTAN=roc_resultCOTAN, 
                                    Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, 
                 Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, 
                 Scanpy=roc_resultScanpy),
                 aes = c("colour"),
            legacy.axes = T,
            linewidth = 1)+theme_light()+
  scale_colour_manual("", 
                      values = methods.color)

 
Five_Clusters_unevenPL <- Five_Clusters + xlab("FPR") + ylab("TPR")

Five_Clusters_unevenPL
```

# Global Summary

## 2 Clusters

```{r,fig.height=10,fig.width=15}
plot_2cl <- ggarrange(TwoClusters_even_nearPL,Two_Clusters_even_mediumPL, Two_Clusters_even_farPL,Two_Clusters_uneven_nearPL, Two_Clusters_uneven_mediumPL,Two_Clusters_uneven_farPL,
          labels = c("Even_Near", "Even_Medium", "Even_Far", "Uneven_Near","Uneven_Medium","Uneven_Far"),
          ncol = 3, nrow = 2, common.legend = T, legend = "bottom")

pdf(paste0(dirOut,"ROC_two_supp.pdf"),width = 15,height = 10)
plot(plot_2cl)
dev.off()

plot_2cl
```

## 3 and 5 Clusters

```{r,fig.height=10,fig.width=10}
plot_3_5 <- ggarrange(Three_Clusters_evenPL,Three_Clusters_unevenPL, NULL, Five_Clusters_unevenPL,
          labels = c("3_Even", "3_Uneven", "", "5_Uneven"),
          ncol = 2, nrow = 2, common.legend = T, legend = "bottom")

pdf(paste0(dirOut,"ROC_three_five_supp.pdf"),width = 10,height = 10)
plot(plot_3_5)
dev.off()

plot_3_5
```

```{r,fig.height=5,fig.width=15}
plot3 <- ggarrange(Two_Clusters_uneven_farPL,Three_Clusters_unevenPL, Five_Clusters_unevenPL,
          labels = c("A", "B", "C"),
          ncol = 3, nrow = 1, common.legend = T, legend = "none")

pdf(paste0(dirOut,"ROC_three_main.pdf"),width = 15,height = 5)
plot(plot3)
dev.off()

plot3
```

## Histogram with gene expression level

```{r}
ground_truth_tot_subset <- as.data.frame(ground_truth_tot[ground_truth_tot$value ==1,])

COTAN_data <- onlyPositive.pVal.Cotan_tot[,c("genes","set_number","clusters","p_values","genesLambda")]

colnames(COTAN_data) <- c("genes","set_number","clusters","p_value_COTAN","genesLambda")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 COTAN_data,
                                 by=c("genes","set_number", "clusters"),
                                 all.x = T, all.y= FALSE)

Seurat_data_base <- deaSeurat_tot_base[,c("genes","set_number","clusters", "p_val")]
colnames(Seurat_data_base) <- c("genes","set_number","clusters","p_value_Seurat")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 Seurat_data_base,
                                 by=c("genes","set_number","clusters"),
                                 all.x = T, all.y= FALSE)

Seurat_data_scTr <- deaSeurat_tot_scTr[,c("genes","set_number","clusters", "p_val")]
colnames(Seurat_data_scTr) <- c("genes","set_number","clusters","p_value_Seurat_scTr")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 Seurat_data_scTr,
                                 by=c("genes","set_number","clusters"),
                                 all.x = T, all.y= FALSE)

Seurat_data_Bimod <- deaSeurat_tot_Bimod[,c("genes","set_number","clusters", "p_val")]
colnames(Seurat_data_Bimod) <- c("genes","set_number","clusters","p_value_Seurat_Bimod")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 Seurat_data_Bimod,
                                 by=c("genes","set_number","clusters"),
                                 all.x = T, all.y= FALSE)


data_Monocle <- deaMonocle_tot[,c("genes","set_number","clusters", "marker_test_p_value")]
colnames(data_Monocle) <- c("genes","set_number","clusters","p_value_Monocle")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 data_Monocle,
                                 by=c("genes","set_number","clusters"),
                                 all.x = T, all.y= FALSE)

data_Scanpy <- deaScanpy_tot[,c("genes","set_number","clusters", "pval")]
colnames(data_Scanpy) <- c("genes","set_number","clusters","p_value_Scanpy")

ground_truth_tot_subset <- merge(ground_truth_tot_subset,
                                 data_Scanpy,
                                 by=c("genes","set_number","clusters"),
                                 all.x = T, all.y= FALSE)

ground_truth_tot_subset <- ground_truth_tot_subset %>% mutate(lambda_bin = cut(genesLambda, breaks=50))

ground_truth_tot_subset[is.na(ground_truth_tot_subset)] <- 1

head(ground_truth_tot_subset)
```

```{r}
bins <- levels(ground_truth_tot_subset$lambda_bin)
bins <- bins[!is.na(bins)]

forBarChart <- as.data.frame(matrix(nrow = length(bins),ncol = 7,data = 0))

colnames(forBarChart) <- c("bins","COTAN","Seurat","Seurat_Bimod","Seurat_scTr","Monocle", "Scanpy")
forBarChart$bins <- bins
rownames(forBarChart) <- forBarChart$bins

#p_val_threshold <- 0.05

for (i in as.vector(forBarChart$bins)) {
  forBarChart[ i,"COTAN"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_COTAN <= cotan.threshold.pval)
  
  forBarChart[ i,"Seurat"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_Seurat <= seurat.threshold.pval)
  
  forBarChart[ i,"Seurat_Bimod"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_Seurat_Bimod <= seurat_bimod.threshold.pval)
  
  forBarChart[ i,"Seurat_scTr"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_Seurat_scTr <= Seurat_scTr.threshold.pval)
  
  forBarChart[ i,"Monocle"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_Monocle <= monocle.threshold.pval)
  
  forBarChart[ i,"Scanpy"] <- sum(ground_truth_tot_subset$lambda_bin == i & ground_truth_tot_subset$p_value_Scanpy <= scanpy.threshold.pval)
  
}

temp.bin <- as.data.frame(str_split(forBarChart$bins,pattern = ",",simplify = T))
temp.bin[,1] <- as.numeric(str_remove(temp.bin[,1],pattern = "[(]"))
temp.bin[,2] <- as.numeric(str_remove(temp.bin[,2],pattern = "[]]"))

forBarChart$bins <- round(rowMeans(temp.bin),digits = 2)


forBarChart <- forBarChart[rowSums(forBarChart[,2:ncol(forBarChart)]) > 0,]

forBarChart.perc <- forBarChart[1:8,]

forBarChart.perc$tot <- rowSums(forBarChart.perc[,2:ncol(forBarChart.perc)])

forBarChart.perc[,2:(ncol(forBarChart.perc)-1)] <- round(forBarChart.perc[,2:(ncol(forBarChart.perc)-1)]/forBarChart.perc$tot,digits = 4)

forBarChart.perc <- pivot_longer(forBarChart.perc,
                               cols = 2:(ncol(forBarChart.perc)-1),
                               names_to = "Methods")

forBarChart.number <- pivot_longer(forBarChart[1:8,],
                               cols = 2:(ncol(forBarChart)),
                               names_to = "Methods")

colnames(forBarChart.number) <- c("bins","Methods", "valueOriginal")

forBarChart.perc <- merge(forBarChart.perc,forBarChart.number, by = c("bins","Methods"))

barChart <- ggplot(forBarChart.perc,aes(x=bins,y=value,fill=Methods))+
  geom_bar(stat="identity")+
  geom_text(aes(label = valueOriginal), 
            position = position_stack(vjust = 0.5),  size=3)+
  scale_fill_manual(
    values = methods.color)+
  theme_light()+
  scale_x_continuous(breaks =forBarChart.perc$bins)+
  theme(axis.title.y=element_blank(),axis.title.x=element_blank(),
      legend.position="none",axis.text.x = element_text(angle = 45, hjust=1)
    )+
  scale_y_continuous(labels = scales::percent_format())

pdf(paste0(dirOut,"EnrichedGenesForExpressionLevel.pdf"),height = 3,width = 5)
barChart
dev.off()

plot(barChart)

```

```{r,fig.width=15,fig.height=5}
plot.with.inset <-
  ggdraw() +
  draw_plot(plot3) +
  draw_plot(as_ggplot(lg), x = 0.18, y = 0.1, width = 0.2, height = .45)+
  draw_plot(barChart, x = 0.78, y = 0.09, width = 0.22, height = .5)

pdf(paste0(dirOut,"Plot3WihtGenesExLevel.pdf"),height = 5,width = 15)
plot(plot.with.inset)
dev.off()

plot(plot.with.inset)
```

## Footer

```{r}
sessionInfo()
```
