---
title: "FDR analysis - Results"
author: "Silvia Giulia Galfrè"
date: "`r Sys.Date()`"
output: html_document
---

## Automatic functions

```{r}
library(COTAN)
library(pROC)
options(parallelly.fork.enable = TRUE)
library(Seurat)
library(monocle3)
library(reticulate)
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)

dirOut <- "Results/FDR/"
if (!dir.exists(dirOut)) {
  dir.create(dirOut)
}

dataSetDir <- "Data/MouseCortexFromLoom/FDR/MergedClusters_For_FDR/"
```

## Cluster Gene Enrichment

```{r}
datasets_csv <- read.csv(file.path(dataSetDir,"Cells_Usage_DataFrame.csv"),
                         row.names = 1
                        ) 

datasets_csv
```

### Define which genes are expressed

For each data set we need to define, independently from the DEA methods, which genes are specific for each cluster. So we need to define first which genes are expressed and which are not expressed. To do so we can take advantage from the fact that we have the original clusters from which the cells were sampled to create the artificial datasets. So looking to the original cluster we define as expressed all genes present in at least the 5% of cells and we define as not expressed the genes completely absent or expressed in less than 3% of cells

```{r}
file.presence <- readRDS("Data/MouseCortexFromLoom/FDR/Results/GenePresence_PerCluster.RDS")

for (file in list.files("Data/MouseCortexFromLoom/SingleClusterRawData/")) {
#  print(file)
  Code <- str_split(file,pattern = "_",simplify = T)[1]
  Time <- str_split(Code,pattern = "e",simplify = T)[2]
  Cluster <- str_split(Code,pattern = "e",simplify = T)[1]
  Cluster <- str_remove(Cluster,pattern = "Cl")
  Cluster <- paste0("E",Time,"-",Cluster)
  file.presence[,Cluster] <- "Absent"
  dataset.cl <- readRDS(file.path("Data/MouseCortexFromLoom/SingleClusterRawData/",
                                         file))
  number.cell.expressing <- rowSums(dataset.cl > 0)
  AbsentThreshold <- round(0.03*dim(dataset.cl)[2],digits = 0)
  PresenceThreshold <- round(0.10*dim(dataset.cl)[2],digits = 0)
  file.presence[names(number.cell.expressing[number.cell.expressing > AbsentThreshold]),Cluster] <- "Uncertain" 
  
  file.presence[names(number.cell.expressing[number.cell.expressing >= PresenceThreshold]),Cluster] <- "Present"
  print(Cluster)
  print(table(file.presence[,Cluster]))
  
  }
```

# 2 Clusters even

## 2_Clusters_even_near

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_near",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan1, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat1, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle1, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy1, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## 2_Clusters_even_medium

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_medium",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan2, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat2, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle2, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy2, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## 2_Clusters_even_far

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_far",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan3, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat3, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle3, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy3, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

# 2 clusters uneven

## 2_Clusters_uneven_near

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_near",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan4, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat4, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle4, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy4, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## 2_Clusters_even_medium

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_medium",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan5, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat5, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle5, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy5, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## 2_Clusters_even_far

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_far",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan6, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat6, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle6, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy6, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

# 3 clusters even

## 3_Clusters_even

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_even",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan7, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat7, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle7, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy7, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## 3_Clusters_uneven

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_uneven",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan8, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat8, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle8, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy8, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

# 5 clusters

## 5_Clusters_uneven

True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "5_Clusters_uneven",] 

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  file.presence.subset <- file.presence[,clusters]
  
  
  #file.presence.subset <- as.matrix(file.presence.subset)
  
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(file.presence.subset),
                                       ncol = ncol(file.presence.subset)))
  rownames(ground_truth) <- rownames(file.presence.subset)
  colnames(ground_truth) <- colnames(file.presence.subset)
  
  ground_truth[file.presence.subset == "Absent"] <- 0
  ground_truth[file.presence.subset == "Present"] <- 1
  ground_truth[file.presence.subset == "Uncertain"] <- 0.5
  
  file.presence.subset <- ground_truth
  
  for (col in colnames(ground_truth)) {
    ground_truth[,col] <- FALSE  
    ground_truth[file.presence.subset[,col] == 1 & rowMeans(file.presence.subset) < 0.65  ,col] <- TRUE
    }
  
  ground_truth$genes <- rownames(ground_truth) 
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind,1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
  
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

### ROC for COTAN

```{r cotan9, echo=TRUE, eval=TRUE}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),method = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)


  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r seurat9, echo=TRUE, eval=TRUE}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r monocle9, echo=TRUE, eval=TRUE}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r scampy9, echo=TRUE, eval=TRUE}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScamPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

```{r}
g2 <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Monolce=roc_resultMonocle, ScamPy=roc_resultScamPy))

g2+ xlab("FPR") + ylab("TPR")
```

## Footer

```{r}
sessionInfo()
```
