---
title: "Results - Log Fold Change"
author: "Marco fantozzi"
format: html
editor: visual
---

## Preamble

```{r}
library(COTAN)
library(pROC)
options(parallelly.fork.enable = TRUE)
#library(Seurat)
#library(monocle3)
#library(reticulate)
library(stringr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)

dirOut <- file.path(".", "Results/FDR/")
if (!dir.exists(dirOut)) {
  dir.create(dirOut)
}

dataSetDir <- file.path("Data/MouseCortexFromLoom/FDR/", "MergedClusters_For_FDR")
```

## Dataset composition

```{r}
datasets_csv <- read.csv(file.path(dataSetDir,"Cells_Usage_DataFrame.csv"),
                         row.names = 1, header = TRUE, quote = '"'
                        ) 

datasets_csv[1:10, ]
```

### Define which genes are expressed

For each data set we need to define, independently from the DEA methods, which genes are specific for each cluster. So we need to define first which genes are expressed and which are not expressed. To do so we can take advantage from the fact that we have the original clusters from which the cells were sampled to create the artificial datasets. So looking to the original cluster we define as expressed all genes present for which there are at least 2 reads every 10 cells. We define as enriched for a cluster the genes that are present and have at least 3 times as many average reads as the average reads outside the cluster.

```{r}
readsLogMeansPerCluster <- readRDS(file.path(dirOut, "readsLogAverageCount_PerCluster.RDS"))

thresholdLFC <- log10(3.0) # presence if 3 times more reads on average
thresholdLFM <- log10(10000 * 2 / 10) # 2 reads each 10 cells

10^thresholdLFC
10^(thresholdLFM-4)
```

# 2 Clusters even

## 2_Clusters_even_near

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_near", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1] 
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_even_near <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_even_nearPL <- TwoClusters_even_near + xlab("FPR") + ylab("TPR")
TwoClusters_even_nearPL
```

## 2_Clusters_even_medium

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_medium", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_even_medium <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_even_mediumPL <- TwoClusters_even_medium + xlab("FPR") + ylab("TPR")
TwoClusters_even_mediumPL
```

## 2_Clusters_even_far

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_even_far", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_even_far <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_even_farPL <- TwoClusters_even_far + xlab("FPR") + ylab("TPR")
TwoClusters_even_farPL
```

# 2 Clusters uneven

## 2_Clusters_uneven_near

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_near", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_uneven_near <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_uneven_nearPL <- TwoClusters_uneven_near + xlab("FPR") + ylab("TPR")
TwoClusters_uneven_nearPL
```

## 2_Clusters_uneven_medium

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_medium", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_uneven_medium <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_uneven_mediumPL <- TwoClusters_uneven_medium + xlab("FPR") + ylab("TPR")
TwoClusters_uneven_mediumPL
```

## 2_Clusters_uneven_far

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "2_Clusters_uneven_far", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
TwoClusters_uneven_far <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

TwoClusters_uneven_farPL <- TwoClusters_uneven_far + xlab("FPR") + ylab("TPR")
TwoClusters_uneven_farPL
```

# 3 clusters

## 3_Clusters_even

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_even", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
ThreeClusters_even <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

ThreeClusters_evenPL <- ThreeClusters_even + xlab("FPR") + ylab("TPR")
ThreeClusters_evenPL
```

## 3_Clusters_uneven

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "3_Clusters_uneven", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
ThreeClusters_uneven <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

ThreeClusters_unevenPL <- ThreeClusters_uneven + xlab("FPR") + ylab("TPR")
ThreeClusters_unevenPL
```

# 5 clusters

## 5_Clusters_uneven

### True vector

```{r}
subset.datasets_csv <-datasets_csv[datasets_csv$Group == "5_Clusters_uneven", ]

ground_truth_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  clusters <- str_split(subset.datasets_csv$Collection[ind], pattern = "_[+]_", simplify = T)
  
  reads.LM.subset <- readsLogMeansPerCluster[, clusters]
  
  ground_truth <- as.data.frame(matrix(nrow = nrow(reads.LM.subset),
                                       ncol = ncol(reads.LM.subset)))
  rownames(ground_truth) <- rownames(reads.LM.subset)
  colnames(ground_truth) <- colnames(reads.LM.subset)

  for (col in 1:ncol(ground_truth)) {
    # log fold change
    ground_truth[, col] <- reads.LM.subset[, col] - rowMeans(reads.LM.subset[, -col, drop = FALSE])
    ground_truth[, col] <- ground_truth[, col] > thresholdLFC & reads.LM.subset[, col] > thresholdLFM
  }

  ground_truth$genes <- rownames(ground_truth)
  ground_truth <- pivot_longer(ground_truth,
                               cols = 1:(ncol(ground_truth)-1),
                               names_to = "clusters")
  ground_truth$data_set <- subset.datasets_csv[ind, 1]
  ground_truth$set_number <- ind 
  ground_truth_tot <- rbind(ground_truth_tot, ground_truth)
}
ground_truth_tot <- ground_truth_tot[2:nrow(ground_truth_tot),]

head(ground_truth_tot)
```

```{r}
length(unique(ground_truth_tot$genes))

sum(ground_truth_tot$value)
```

### ROC for COTAN

```{r}
onlyPositive.pVal.Cotan_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",
                      subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  deaCOTAN <- getClusterizationData(dataset,clName = "mergedClusters")[[2]]
  pvalCOTAN <- pValueFromDEA(deaCOTAN,
              numCells = getNumCells(dataset),adjustmentMethod = "none")
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.names <- c(cl.names,
                  str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1])
    cl.names <- cl.names[!is.na(cl.names)]
  }
  colnames(deaCOTAN) <- cl.names
  colnames(pvalCOTAN) <- cl.names
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  onlyPositive.pVal.Cotan <- pvalCOTAN

    for (cl in cl.names) {
    print(cl)
    #temp.DEA.CotanSign <- deaCOTAN[rownames(pvalCOTAN[pvalCOTAN[,cl] < 0.05,]) ,]
    onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl] <-  1 #onlyPositive.pVal.Cotan[rownames(deaCOTAN[deaCOTAN[,cl] < 0,]),cl]+1
    }
    
    onlyPositive.pVal.Cotan$genes <- rownames(onlyPositive.pVal.Cotan) 
    onlyPositive.pVal.Cotan <- pivot_longer(onlyPositive.pVal.Cotan,
                                            values_to = "p_values",
                               cols = 1:(ncol(onlyPositive.pVal.Cotan)-1),
                               names_to = "clusters")
  onlyPositive.pVal.Cotan$data_set <- subset.datasets_csv[ind,1]
  onlyPositive.pVal.Cotan$set_number <- ind 
  onlyPositive.pVal.Cotan_tot <- rbind(onlyPositive.pVal.Cotan_tot, onlyPositive.pVal.Cotan)
  
  }
```

```{r}
onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[2:nrow(onlyPositive.pVal.Cotan_tot),]

onlyPositive.pVal.Cotan_tot <- merge.data.frame(onlyPositive.pVal.Cotan_tot,
                                                ground_truth_tot,by = c("genes","clusters","data_set","set_number"),all.x = T,all.y = F)

#onlyPositive.pVal.Cotan_tot <- onlyPositive.pVal.Cotan_tot[order(onlyPositive.pVal.Cotan_tot$p_values,
 #                                                                decreasing = F),]
# df <- as.data.frame(matrix(nrow = nrow(onlyPositive.pVal.Cotan_tot),ncol = 3)) 
# colnames(df) <- c("TPR","FPR","Method")
# df$Method <- "COTAN"
# 
# Positive <- sum(onlyPositive.pVal.Cotan_tot$value) 
# Negative <- sum(!onlyPositive.pVal.Cotan_tot$value)
# 
# for (i in 1:nrow(onlyPositive.pVal.Cotan_tot)) {
#   df[i,"TPR"] <- sum(onlyPositive.pVal.Cotan_tot[1:i,"value"])/Positive
#   df[i,"FPR"] <- (i-sum(onlyPositive.pVal.Cotan_tot[1:i,"value"]))/Negative
  
# Convert TRUE/FALSE to 1/0
onlyPositive.pVal.Cotan_tot$value <- as.numeric(onlyPositive.pVal.Cotan_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultCOTAN <- roc(onlyPositive.pVal.Cotan_tot$value, 1 - onlyPositive.pVal.Cotan_tot$p_values)

# Plot the ROC curve
#plot(roc_resultCOTAN)
```

### ROC for Seurat

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat scTransform

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_ScTransform_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_scTr <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Seurat Bimod

```{r}
deaSeurat_tot <- NA

for (ind in 1:dim(subset.datasets_csv)[1]) {
  #print(ind)
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  
  deaSeurat <- read.csv(file.path(dirOut,paste0(file.code,"Seurat_DEA_Bimod_genes.csv")), row.names = 1)
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaSeurat[deaSeurat$cluster == cl.val,]$cluster <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "cluster",
                                     replacement = "clusters") 
  colnames(deaSeurat) <- str_replace(colnames(deaSeurat),
                                     pattern = "gene",
                                     replacement = "genes") 
  
  deaSeurat$data_set <- subset.datasets_csv[ind,1]
  deaSeurat$set_number <- ind 
  deaSeurat_tot <- rbind(deaSeurat_tot, deaSeurat)
  
  
  }
 deaSeurat_tot <-  deaSeurat_tot[2:nrow(deaSeurat_tot),]
 
 deaSeurat_tot <- merge.data.frame(deaSeurat_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaSeurat_tot$value <- as.numeric(deaSeurat_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultSeurat_Bimod <- roc(deaSeurat_tot$value, 1 - deaSeurat_tot$p_val)

# Plot the ROC curve
#plot(roc_resultSeurat)
```

### ROC for Monocle

```{r}
deaMonocle_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))
  
  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaMonocle <- read.csv(file.path(dirOut,paste0(file.code,"Monocle_DEA_genes.csv")),row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaMonocle[deaMonocle$cell_group == cl.val,"cell_group"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)
  
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "cell_group",
                                     replacement = "clusters") 
  colnames(deaMonocle) <- str_replace(colnames(deaMonocle),
                                     pattern = "gene_id",
                                     replacement = "genes")
  
  deaMonocle$data_set <- subset.datasets_csv[ind,1]
  deaMonocle$set_number <- ind 
  deaMonocle <- as.data.frame(deaMonocle)
  deaMonocle_tot <- rbind(deaMonocle_tot, deaMonocle)
  
  
  }
 deaMonocle_tot <-  deaMonocle_tot[2:nrow(deaMonocle_tot),]
 
 deaMonocle_tot <- merge.data.frame(deaMonocle_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaMonocle_tot$value <- as.numeric(deaMonocle_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultMonocle <- roc(deaMonocle_tot$value, 1 - deaMonocle_tot$marker_test_p_value)

# Plot the ROC curve
#plot(roc_resultMonocle)
```

### ROC from ScamPy

```{r}
deaScamPy_tot <- NA
for (ind in 1:dim(subset.datasets_csv)[1]) {
  file.code <- paste0(subset.datasets_csv$Group[ind],"_",subset.datasets_csv$Collection[ind])
  dataset <- readRDS(file = file.path(dataSetDir,paste0(file.code,".RDS")))

  clusterization <- getClusterizationData(dataset, clName = "mergedClusters")[[1]]
  #print(file.code)
  
  deaScamPy <- read.csv(file.path(dirOut,paste0(file.code,"ScanPy_DEA_genes.csv")),
                        row.names = 1) 
  
  cl.names <- NA
  
  for (cl.val in  unique(clusterization)) {
    #print(cl.val)
    cl.name <- str_split(names(clusterization[clusterization == cl.val])[1],
                            pattern = "_",simplify = T)[1]
    cl.names <- c(cl.names,cl.name)
    cl.names <- cl.names[!is.na(cl.names)]
  
    deaScamPy[deaScamPy$clusters == paste0("cl",cl.val),"clusters"] <- cl.name   
  
  }
  
  clusters <- str_split(datasets_csv$Collection[ind],pattern = "_[+]_",simplify = T)


  deaScamPy$data_set <- subset.datasets_csv[ind,1]
  deaScamPy$set_number <- ind 
  deaScamPy_tot <- rbind(deaScamPy_tot, deaScamPy)
  
  
  }
 deaScamPy_tot <-  deaScamPy_tot[2:nrow(deaScamPy_tot),]
 
 deaScamPy_tot <- merge.data.frame(deaScamPy_tot,
                                   ground_truth_tot,
                                   by = c("genes","clusters","data_set","set_number"),
                                   all.x = T,all.y = F)

# Convert TRUE/FALSE to 1/0
deaScamPy_tot$value <- as.numeric(deaScamPy_tot$value)

# Compute the ROC curve - note that we invert the p-values with 1 - p_values
roc_resultScamPy <- roc(deaScamPy_tot$value, 1 - deaScamPy_tot$pval)

# Plot the ROC curve
#plot(roc_resultScamPy)
```

### Summary ROC for all methods

```{r}
FiveClusters_uneven <- ggroc(list(COTAN=roc_resultCOTAN, Seurat=roc_resultSeurat,
                 Seurat_scTr = roc_resultSeurat_scTr, Seurat_Bimod = roc_resultSeurat_Bimod,
                 Monocle=roc_resultMonocle, ScamPy=roc_resultScamPy))

FiveClusters_unevenPL <- FiveClusters_uneven + xlab("FPR") + ylab("TPR")
FiveClusters_unevenPL
```

# Global Summary

## 2 Clusters

```{r,fig.height=10,fig.width=15}
ggarrange(TwoClusters_even_nearPL,TwoClusters_even_mediumPL, TwoClusters_even_farPL,TwoClusters_uneven_nearPL, TwoClusters_uneven_mediumPL,TwoClusters_uneven_farPL,
          labels = c("Even_Near", "Even_Medium", "Even_Far", "Uneven_Near","Uneven_Medium","Uneven_Far"),
          ncol = 3, nrow = 2, common.legend = T, legend = "bottom")
```

## 3 and 5 Clusters

```{r,fig.height=10,fig.width=10}
ggarrange(ThreeClusters_evenPL,ThreeClusters_unevenPL, NULL, FiveClusters_unevenPL,
          labels = c("3_Even", "3_Uneven", "", "5_Uneven"),
          ncol = 2, nrow = 2, common.legend = T, legend = "bottom")
```

## Footer

```{r}
sessionInfo()
```