---
title: "General Coex Evaluation"
---

```{r}
library(tidyr)
library(stringr)
library(ggplot2)
library(gghalves)
library(ggstatsplot)
library("gridExtra")
library(cowplot)
library(Matrix)
library(ggstance)
library(scales)
#library(COTAN)
```

```{r}
genesList <- list(
  "NPGs"= 
    c("Nes", "Vim", "Sox2", "Sox1", "Notch1", "Hes1", "Hes5", "Pax6"),
  "PNGs"= 
    c("Map2", "Tubb3", "Neurod1", "Nefm", "Nefl", "Dcx", "Tbr1"),
  "hk"= 
    c("Calm1", "Cox6b1", "Ppia", "Rpl18", "Cox7c", "Erh", "H3f3a",
      "Taf1", "Taf2", "Gapdh", "Actb", "Golph3", "Zfr", "Sub1",
      "Tars", "Amacr"),
  "layers" = 
    c("Reln","Lhx5","Cux1","Satb2","Tle1","Mef2c","Rorb","Sox5","Bcl11b","Fezf2","Foxp2","Ntf3","Rasgrf2","Pvrl3", "Cux2","Slc17a6", "Sema3c","Thsd7a", "Sulf2", "Kcnk2","Grik3", "Etv1", "Tle4", "Tmem200a", "Glra2", "Etv1","Htr1f", "Sulf1","Rxfp1", "Syt6") 
  # From https://www.science.org/doi/10.1126/science.aam8999
)
```

## COEX and Correlations

Import COTAN coex and correlations data from file

```{r}
table.tot.hk <- NA
table.tot.neural <- NA
for (file in list.files("CoexData/",pattern = "Corr|Coex")) {
  corr <- readRDS(paste0("CoexData/",file))
  corr <- as.data.frame(as.matrix(corr))
  if(str_detect(file,pattern = "CotanCoex")){
    code <- "COTAN coex"
  }else if(str_detect(file,pattern = "CorrSCT")){
    code <- "Seurat SCT corr."
  }else if(str_detect(file,pattern = "Corr")){
    code <- "Seurat corr."
  }
  
  
  table.hk <- as.data.frame(corr[,genesList$hk])
  table.hk$Gene1 <- rownames(table.hk)
  
  table.hk <- pivot_longer(as.data.frame(table.hk),cols = c(1:(ncol(table.hk)-1)),names_to = "Gene2")
  table.hk$Method <- code

  table.tot.hk <- rbind(table.tot.hk,table.hk)
  table.tot.hk <- table.tot.hk[! table.tot.hk$Gene1 == table.tot.hk$Gene2,]
 
  #Not hk
  table.neural <- as.data.frame(corr[rownames(corr) %in% c(genesList$NPGs,genesList$PNGs,genesList$layers),colnames(corr) %in% c(genesList$NPGs,genesList$PNGs,genesList$layers)])
  table.neural$Gene1 <- rownames(table.neural)
  
  table.neural <- pivot_longer(as.data.frame(table.neural),cols = c(1:(ncol(table.neural)-1)),names_to = "Gene2")
  table.neural$Method <- code
  
  table.tot.neural <- rbind(table.tot.neural,table.neural)
  table.tot.neural <- table.tot.neural[! table.tot.neural$Gene1 == table.tot.neural$Gene2,]
}
```

```{r}
table.tot.hk <- table.tot.hk[2:nrow(table.tot.hk),]
table.tot.hk$GeneType <- "Constitutive" 
table.tot.neural <- table.tot.neural[2:nrow(table.tot.neural),]
table.tot.neural$GeneType <- "Neural"

## For ECDL
tmp.tot <- NA
for (method in unique(table.tot.hk$Method)) {
  tmp <- as.data.frame(table.tot.hk[table.tot.hk$Method == method,])
  tmp$n <- NA
  tmp[order(abs(tmp$value)),]$n <- c(1:dim(tmp)[1])/(dim(tmp)[1]+1)
  tmp$Y_axis <- 1-tmp$n
  tmp.tot <- rbind(tmp.tot,tmp)
}
tmp.tot <- tmp.tot[2:nrow(tmp.tot),]
table.tot.hk <- tmp.tot

tmp.tot <- NA
for (method in unique(table.tot.neural$Method)) {
  tmp <- as.data.frame(table.tot.neural[table.tot.neural$Method == method,])
  tmp$n <- NA
  tmp[order(abs(tmp$value)),]$n <- c(1:dim(tmp)[1])/(dim(tmp)[1]+1)
  tmp$Y_axis <- 1-tmp$n
  tmp.tot <- rbind(tmp.tot,tmp)
}
tmp.tot <- tmp.tot[2:nrow(tmp.tot),]
table.tot.neural <- tmp.tot

############

table.tot <- rbind(table.tot.hk,table.tot.neural) 
table.tot <- table.tot[!table.tot$Gene1 == table.tot$Gene2,]

identical(table.tot[table.tot$Method == "COTAN coex",]$Gene1,
          table.tot[table.tot$Method == "Seurat corr.",]$Gene1 )
identical(table.tot[table.tot$Method == "COTAN coex",]$Gene2,
          table.tot[table.tot$Method == "Seurat corr.",]$Gene2 )

head(table.tot)
```

```{r}
table.tot$ValAbs <- abs(table.tot$value)
table.tot$LogValAbs <- log(table.tot$ValAbs+0.001)
table.tot$SqrValAbs <- table.tot$ValAbs**2

ggplot(table.tot,aes(x=Method,y=LogValAbs, fill=Method)) +
geom_half_violin(alpha=0.8) +
  #geom_point(position = "jitter", size= 0.1,alpha = 0.3)+ 
  geom_half_boxplot(width=0.1, alpha=0.8,side = "r") +
  theme_bw()+   theme(legend.position="none")+#ylim(0,1)+
  facet_grid(GeneType ~ . , space = "free_y", scales = "free")+
  scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))

```

```{r, warning=FALSE}
ggplot(table.tot[table.tot$GeneType == "Neural",], aes(x = ValAbs, y = -0.5)) +

  # horizontal box plot
  geom_boxploth(aes(fill = Method)) +

  # normal density plot
  geom_density(aes(x = ValAbs), inherit.aes = FALSE) +

  # # vertical lines at Q1 / Q2 / Q3
  # stat_boxploth(geom = "vline", aes(xintercept = ..xlower..)) +
   stat_boxploth(geom = "vline", aes(xintercept = ..xmiddle..)) +
  # stat_boxploth(geom = "vline", aes(xintercept = ..xupper..)) +
  # 
   facet_grid(Method ~ .) +
  
  # reproduce original chart's color scale (o/w ordered factors will result
  # in viridis scale by default, using the current version of ggplot2)
  scale_fill_discrete()+scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()+   theme(legend.position="none")
```

In this case we don't see differences between coex and correlations.

```{r, warning=FALSE}
ggplot(table.tot[table.tot$GeneType == "Constitutive",], aes(x = ValAbs, y = -0.5)) +

  # horizontal box plot
  geom_boxploth(aes(fill = Method), outlier.size = 0.5,width=3) +

  # normal density plot
  geom_density(aes(x = ValAbs), inherit.aes = FALSE) +

  # # vertical lines at Q1 / Q2 / Q3
  # stat_boxploth(geom = "vline", aes(xintercept = ..xlower..)) +
   stat_boxploth(geom = "vline", aes(xintercept = ..xmiddle..)) +
  # stat_boxploth(geom = "vline", aes(xintercept = ..xupper..)) +
  # 
   facet_grid(Method ~ .) +
  
  # reproduce original chart's color scale (o/w ordered factors will result
  # in viridis scale by default, using the current version of ggplot2)
  scale_fill_discrete()+scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()+   theme(legend.position="none")
```

We can clearly observe that the coex is generally more near to zero than the correlations.

```{r}

ggplot(table.tot,aes(x=ValAbs, color=Method, y = Y_axis)) + 
  geom_point(alpha=0.8) +
  scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  scale_y_continuous(trans= "log",breaks = c(0,0.1,0.25,0.5)
                         )+
  scale_x_continuous(trans= "log",
                     breaks = c(0.1,0.2,0.3, 0.4,0.5,0.8), limits = c(0.01,0.85))+
  theme_bw()+   theme(legend.position="bottom")+ 
  facet_grid(GeneType ~ . )
```

```{r}
ggplot(table.tot.hk,aes(x=abs(value), color=Method))+stat_ecdf(geom = "point", size = 0.5 )+scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  scale_y_continuous(trans= "exp")
```

```{r}

ggplot(table.tot,aes(x=Method,y=ValAbs, fill=Method)) +
#geom_half_boxplot(alpha=0.8) +
  #geom_point(position = "jitter", size= 0.1,alpha = 0.3)+ 
  geom_boxplot(alpha=0.8,outlier.size = 0.1,outlier.colour='blue', alpha=0.1) +
  theme_bw()+   theme(legend.position="none")+ylim(0,0.2)+
  facet_grid(GeneType ~ . , space = "free_y", scales = "free")+
  scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))
```

```{r}
ggplot(table.tot,aes(x=Method,y=ValAbs, fill=GeneType)) +
  geom_boxplot(alpha=0.8) +
  #geom_hline(yintercept = -3.39)+
  #geom_hline(yintercept = -1.86)+
  #geom_point(position = "jitter", size= 0.1,alpha = 0.3)+ 
  #geom_half_boxplot(width=0.1, alpha=0.8,side = "r") +
  theme_bw()+   #theme(legend.position="none")+
  scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))

```

## *p*-values

```{r}
table.tot.hk <- NA
table.tot.neural <- NA
for (file in list.files("CoexData/",pattern = "PValue")) {
  pval <- readRDS(paste0("CoexData/",file))
  #pval <- as.data.frame(as.matrix(pval))
  if(str_detect(file,pattern = "Cotan")){
    code <- "COTAN p-values"
    if(all(diag(pval)) == 1){
      
      diag(pval) = 0
    }
  }else if(str_detect(file,pattern = "PValuesSCT")){
    code <- "Seurat SCT p-values"
  }else if(str_detect(file,pattern = "SeuratPValues")){
    code <- "Seurat p-values"
  }
  
  
  table.hk <- as.data.frame(pval[,genesList$hk])
  table.hk$Gene1 <- rownames(table.hk)
  
  table.hk <- pivot_longer(as.data.frame(table.hk),cols = c(1:(ncol(table.hk)-1)),names_to = "Gene2")
  table.hk$Method <- code

  table.tot.hk <- rbind(table.tot.hk,table.hk)
  table.tot.hk <- table.tot.hk[! table.tot.hk$Gene1 == table.tot.hk$Gene2,]
 
  #Not hk
  table.neural <- as.data.frame(pval[rownames(pval) %in% c(genesList$NPGs,genesList$PNGs,genesList$layers),colnames(pval) %in% c(genesList$NPGs,genesList$PNGs,genesList$layers)])
  table.neural$Gene1 <- rownames(table.neural)
  
  table.neural <- pivot_longer(as.data.frame(table.neural),cols = c(1:(ncol(table.neural)-1)),names_to = "Gene2")
  table.neural$Method <- code
  
  table.tot.neural <- rbind(table.tot.neural,table.neural)
  table.tot.neural <- table.tot.neural[! table.tot.neural$Gene1 == table.tot.neural$Gene2,]
}

table.tot.hk <- table.tot.hk[2:nrow(table.tot.hk),]
table.tot.hk$GeneType <- "Constitutive" 
table.tot.neural <- table.tot.neural[2:nrow(table.tot.neural),]
table.tot.neural$GeneType <- "Neural"

## For ECDL
tmp.tot <- NA
for (method in unique(table.tot.hk$Method)) {
  tmp <- as.data.frame(table.tot.hk[table.tot.hk$Method == method,])
  tmp$n <- NA
  tmp[order(abs(tmp$value)),]$n <- c(1:dim(tmp)[1])/(dim(tmp)[1]+1)
  tmp$Y_axis <- 1-tmp$n
  tmp.tot <- rbind(tmp.tot,tmp)
}
tmp.tot <- tmp.tot[2:nrow(tmp.tot),]
table.tot.hk <- tmp.tot

tmp.tot <- NA
for (method in unique(table.tot.neural$Method)) {
  tmp <- as.data.frame(table.tot.neural[table.tot.neural$Method == method,])
  tmp$n <- NA
  tmp[order(abs(tmp$value)),]$n <- c(1:dim(tmp)[1])/(dim(tmp)[1]+1)
  tmp$Y_axis <- 1-tmp$n
  tmp.tot <- rbind(tmp.tot,tmp)
}
tmp.tot <- tmp.tot[2:nrow(tmp.tot),]
table.tot.neural <- tmp.tot

############

table.tot <- rbind(table.tot.hk,table.tot.neural) 
table.tot <- table.tot[!table.tot$Gene1 == table.tot$Gene2,]

identical(table.tot[table.tot$Method == "COTAN p-val",]$Gene1,
          table.tot[table.tot$Method == "Seurat p-val.",]$Gene1 )


head(table.tot)
```

### Constitutive genes

```{r}
ggplot(table.tot[table.tot$GeneType == "Constitutive",], aes(y = value)) +
  geom_boxplot(aes(fill = Method), outlier.size = 0.5,width=3) +
  scale_fill_discrete()+scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()#+   theme(legend.position="none")
```

```{r}
ggplot(table.tot.hk,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+
  scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  theme_bw()
```

```{r}
ggplot(table.tot.hk,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+ scale_x_continuous(trans = log_trans(base = 10),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),limits = c(10^(-100),0.05)) + ylim(0,0.4) +scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  theme_bw()
```

We can observe from this plot that COTAN has lower false positive error in fact the number of constitutive genes with *p-*values lower than 0.05 is quite smaller than from Seurat (both normalization).

```{r}
ggplot(table.tot[table.tot$GeneType == "Constitutive",],aes(x=value, color=Method))+geom_density()+xlim(0,0.05)+
  scale_color_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw() 
```

### Neural genes

```{r}
ggplot(table.tot[table.tot$GeneType == "Neural",], aes(y = (value))) +
  geom_boxplot(aes(fill = Method), outlier.size = 0.5,width=3) +
  scale_fill_discrete()+scale_fill_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw() #+ theme(legend.position="none")
```

```{r}
ggplot(table.tot.neural,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+scale_color_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()
```

```{r}
ggplot(table.tot.neural,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+ 
  scale_x_continuous(trans = log_trans(base = 10),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(0.0001,0.05)) +
  scale_color_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()
```

```{r}
ggplot(table.tot.neural,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+ 
  scale_x_continuous(trans = log_trans(base = 10),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(10^(-300),0.05)) + ylim(0,0.5)+
  scale_color_manual(values = c("#8856A7","#EDF8FB","#B3CDE3" ))+
  theme_bw()
```

### Direct comparision

```{r}
ggplot(table.tot,aes(x=value, color=Method))+geom_density()+
  facet_grid(GeneType ~ .,scales="free")+
  scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  theme_bw()
```

```{r}
ggplot(table.tot,aes(x=value,y=n, color=Method))+geom_point(size = 0.5 )+ 
  scale_x_continuous(trans = log_trans(base = 10),
    breaks = trans_breaks("log10", function(x) 10^x),
    labels = trans_format("log10", math_format(10^.x)),
    limits = c(0.0001,0.05)) + 
  facet_grid(. ~ GeneType ,scales="free") +
  scale_color_manual(values = c("#8856A7","gray","#B3CDE3" ))+
  theme_bw()+ theme(legend.position="bottom")
```

------------------------------------------------------------------------

```{r}
Sys.time()
```

```{r}
sessionInfo()

```
