---
title: "Not Uniform Transcriptomic clusters with Logistic Regression"
author: "Silvia Galfrè"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preamble

```{r preamble}
library(ggplot2)
library(zeallot)
library(data.table)
library(parallelDist)
library(tidyr)
library(tidyverse)
library(caret)
theme_set(theme_bw())
library(COTAN)
library(stringr)
library(nnet)

options(parallelly.fork.enable = TRUE)

inDir <- file.path("Data/MouseCortexFromLoom/")

#setLoggingLevel(2)
#setLoggingFile(file.path(inDir, "MixingClustersGDI_ForebrainDorsal.log"))

outDir <- file.path("Data/MouseCortexFromLoom/NotCTSingleClusters/")
if (!file.exists(outDir)) {
  dir.create(outDir)
}


result.df <- data.frame()
```

In this part we want to show that if COTAN detects more division for cluster defined by other methods, this division is consitent and real. To test it we will use a simple logistic regression. 


## Cluster 525 e15.0



```{r, eval=FALSE, echo=TRUE}
e15.0 <- readRDS(paste0(inDir,"e15.0_ForebrainDorsal.cotan.RDS")
cells <- getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]][getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]] == "525"]

cells <-names(cells)

write.csv(getRawData(e15.0)[,cells],paste0(outDir,"cl525/RawDataCL525.csv"))

cl525 <-  automaticCOTANObjectCreation(raw = getRawData(e15.0)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl525",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)

```

```{r, eval=TRUE,echo=FALSE}
cl525 <- readRDS(paste0(outDir,"cl525.cotan.RDS"))

cl <- "Cl525E15.0"
n.cells <- getNumCells(cl525)
n.cls <- length(unique(getClusters(cl525)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
colnames(result.df) <- c("OriginalCluster","CellNumber","COTANClusterNumber","Accuracy" )
result.df
```


```{r}
cl525GDI <- calculateGDI(cl525)

subsetGDICl525 <- cl525GDI[cl525GDI$sum.raw.norm > 7,]
top.GDI.genes <- rownames(subsetGDICl525[order(subsetGDICl525$GDI,decreasing = T),])[1:50]

GDIPlot(cl525,genes = "",GDIIn = cl525GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl525, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl525 <- addClusterization(cl525, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl525, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl525 <- addClusterization(cl525, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl525,paste0(outDir,"cl525.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```

So there isn't a perfect separation due to the cell bar code.


```{r}
cl525ClsMarkers <- findClustersMarkers(cl525,n = 200,method = "BH")

head(cl525ClsMarkers)
```

```{r, eval=FALSE, echo=FALSE}
write.csv(cl525ClsMarkers,paste0(outDir,"cl525/ClustersMarkers.csv"))
write.csv(getClusterizationData(cl525)[[1]],file = paste0(outDir,"cl525/CellClusters.csv"))
```


```{r}
data <- getNormalizedData(cl525)

data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl525)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
train.data$COTAN_Cl.code <- as.factor(train.data$COTAN_Cl.code)
test.data$COTAN_Cl.code <- as.factor(test.data$COTAN_Cl.code)
# Fit the model
#model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 75))
model <- multinom(COTAN_Cl.code ~ ., data = train.data, maxit = 500)
probabilities <- predict(model, newdata = test.data, type = "probs")

# Find the class with the highest probability for each case
predicted.classes <- apply(probabilities, 1, which.max)

# Adjust predicted classes to match your factor levels
levels <- levels(train.data$COTAN_Cl.code)
predicted.classes <- levels[predicted.classes]

# Calculate model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy
print(accuracy)
```
So also a simple logistic regression confirm that this cluster is in fact formed by four different cell populations.

Now lets try to search the identity of these cells.

Since COTAN can detect also very lowly expressed genes, but all the literature methods not, the online methods are bias for highly expressed genes. So, we will take all genes detected as enriched by COTAN and filter also for the highly expressed ones.


```{r}
top <- cl525ClsMarkers[cl525ClsMarkers$Gene %in% rownames(cl525GDI[cl525GDI$sum.raw.norm > 7,]) & cl525ClsMarkers$Score > 0,]
top
```
```{r}
write.csv(top,paste0(outDir,"cl525/TopClustersMarkers.csv"))
```



## Cluster 511 e15.0

```{r, eval=FALSE, echo=TRUE}
cells <- getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]][getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]] == "511"]

cells <-names(cells)

cl511 <-  automaticCOTANObjectCreation(raw = getRawData(e15.0)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl511",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)

```
```{r, eval=TRUE,echo=FALSE}
cl511 <- readRDS(paste0(outDir,"cl511.cotan.RDS"))

cl <- "Cl511E15.0"
n.cells <- getNumCells(cl511)
n.cls <- length(unique(getClusters(cl511)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
result.df
```

```{r}
cl511GDI <- calculateGDI(cl511)

subsetGDIcl511 <- cl511GDI[cl511GDI$sum.raw.norm > 7,]
top.GDI.genes <- rownames(subsetGDIcl511[order(subsetGDIcl511$GDI,decreasing = T),])[1:50]

GDIPlot(cl511,genes = "",GDIIn = cl511GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl511, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl511 <- addClusterization(cl511, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl511, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl511 <- addClusterization(cl511, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl511,paste0(outDir,"cl511.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```




```{r}
data <- getNormalizedData(cl511)
data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl511)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)-1

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
# Fit the model
model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 50))
# Summarize the model
summary(model)
# Make predictions
probabilities <- model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy

accuracy
```

So also in this case a logistic regression confirm the difference in cell population detected by COTAN.

```{r}
cl511ClsMarkers <- findClustersMarkers(cl511,n = 200,method = "BH")
cl511ClsMarkers
```

```{r, eval=FALSE, echo=FALSE}
write.csv(cl511ClsMarkers,paste0(outDir,"cl511/ClustersMarkers.csv"))

write.csv(getClusterizationData(cl511)[[1]],file = paste0(outDir,"cl511/CellClusters.csv"))
```

## Cluster 184 e15.0

```{r, eval=TRUE,echo=FALSE}
cl184 <- readRDS(paste0(outDir,"cl184.cotan.RDS"))
cl <- "Cl184E15.0"
n.cells <- getNumCells(cl184)
n.cls <- length(unique(getClusters(cl184)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
result.df
```


```{r, eval=FALSE, echo=TRUE}
cells <- getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]][getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]] == "184"]

cells <-names(cells)

cl184 <-  automaticCOTANObjectCreation(raw = getRawData(e15.0)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl184",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)

```


```{r}
cl184GDI <- calculateGDI(cl184)

subsetGDIcl184 <- cl184GDI[cl184GDI$sum.raw.norm > 7,]
top.GDI.genes <- rownames(subsetGDIcl184[order(subsetGDIcl184$GDI,decreasing = T),])[1:50]

GDIPlot(cl184,genes = "",GDIIn = cl184GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl184, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl184 <- addClusterization(cl184, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl184, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl184 <- addClusterization(cl184, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl184,paste0(outDir,"cl184.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```



```{r}
data <- getNormalizedData(cl184)
data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl184)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)-1

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
# Fit the model
model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 50))
# Summarize the model
summary(model)
# Make predictions
probabilities <- model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy

accuracy

```

So also in this case a logistic regression confirm the difference in cell population detected by COTAN.

```{r}
cl184ClsMarkers <- findClustersMarkers(cl184,n = 200,method = "BH")
cl184ClsMarkers
```

```{r, echo=FALSE, eval=FALSE}
write.csv(cl184ClsMarkers,paste0(outDir,"cl184/ClustersMarkers.csv"))
write.csv(getClusterizationData(cl184)[[1]],file = paste0(outDir,"cl184/CellClusters.csv"))
```

## Cluster 432 e13.5

```{r, eval=TRUE,echo=FALSE}
cl432  <- readRDS(paste0(outDir,"cl432.cotan.RDS"))
cl <- "Cl432E13.5"
n.cells <- getNumCells(cl432)
n.cls <- length(unique(getClusters(cl432)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
result.df

```


```{r, echo=TRUE, eval=FALSE}
e13.5 <- readRDS(paste0(inDir,"e13.5_ForebrainDorsal.cotan.RDS"))

cells <- getClusterizationData(e13.5,clName = "original.clusters")[[1]][getClusterizationData(e13.5,clName = "original.clusters")[[1]] == "432"]

cells <-names(cells)

cl432 <-  automaticCOTANObjectCreation(raw = getRawData(e13.5)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl432",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)
```

```{r}
cl432GDI <- calculateGDI(cl432)

subsetGDIcl432 <- cl432GDI[cl432GDI$sum.raw.norm > 5,]
top.GDI.genes <- rownames(subsetGDIcl432[order(subsetGDIcl432$GDI,decreasing = T),])[1:25]

GDIPlot(cl432,genes = list("topGDI"=top.GDI.genes),GDIIn = cl432GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl432, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl432 <- addClusterization(cl432, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl432, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl432 <- addClusterization(cl432, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl432,paste0(outDir,"cl432.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```



```{r}
data <- getNormalizedData(cl432)
data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl432)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)-1

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
# Fit the model
model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 50))
# Summarize the model
summary(model)
# Make predictions
probabilities <- model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy

accuracy

```

So also in this case a logistic regression confirm the difference in cell population detected by COTAN.

```{r}
cl432ClsMarkers <- findClustersMarkers(cl432,n = 200,method = "BH")
cl432ClsMarkers
```

```{r, echo=FALSE, eval=FALSE}
write.csv(cl432ClsMarkers,paste0(outDir,"cl432/ClustersMarkers.csv"))
write.csv(getClusterizationData(cl432)[[1]],file = paste0(outDir,"cl432/CellClusters.csv"))
```



## Cluster 510 e15.0

```{r, eval=TRUE,echo=FALSE}
cl510e15.0  <- readRDS(paste0(outDir,"cl510e15.0.cotan.RDS"))

cl <- "Cl510E15.0"
n.cells <- getNumCells(cl510e15.0 )
n.cls <- length(unique(getClusters(cl510e15.0)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
result.df
```


```{r,eval=FALSE, echo=TRUE}
cells <- getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]][getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]] == "510"]

cells <-names(cells)

cl510e15.0 <-  automaticCOTANObjectCreation(raw = getRawData(e15.0)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl510e15.0",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)

```

```{r}
cl510e15.0GDI <- calculateGDI(cl510e15.0)

subsetGDIcl510e15.0 <- cl510e15.0GDI[cl510e15.0GDI$sum.raw.norm > 5,]
top.GDI.genes <- rownames(subsetGDIcl510e15.0[order(subsetGDIcl510e15.0$GDI,decreasing = T),])[1:25]

GDIPlot(cl510e15.0,genes = list("topGDI"=top.GDI.genes),GDIIn = cl510e15.0GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl510e15.0, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl510e15.0 <- addClusterization(cl510e15.0, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl510e15.0, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl510e15.0 <- addClusterization(cl510e15.0, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl510e15.0,paste0(outDir,"cl510e15.0.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```



```{r}
data <- getNormalizedData(cl510e15.0)
data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl510e15.0)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)-1

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
# Fit the model
model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 50))
# Summarize the model
summary(model)
# Make predictions
probabilities <- model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy

accuracy

```

So also in this case a logistic regression confirm the difference in cell population detected by COTAN.

```{r}
cl510e15.0ClsMarkers <- findClustersMarkers(cl510e15.0,n = 200,method = "BH")
cl510e15.0ClsMarkers
```

```{r, echo=FALSE, eval=FALSE}
write.csv(cl510e15.0ClsMarkers,paste0(outDir,"cl510e15.0/ClustersMarkers.csv"))
write.csv(getClusterizationData(cl510e15.0)[[1]],file = paste0(outDir,"cl510e15.0/CellClusters.csv"))
```

## Cluster 511 e15.0

```{r, eval=TRUE,echo=FALSE}
cl511e15.0  <- readRDS(paste0(outDir,"cl511e15.0.cotan.RDS"))

cl <- "Cl511E15.0"
n.cells <- getNumCells(cl511e15.0)
n.cls <- length(unique(getClusters(cl511e15.0)))

result.df <- rbind(result.df,c(cl,n.cells,n.cls,NA))
result.df
```



```{r,eval=FALSE, echo=TRUE}
cells <- getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]][getClusterizationData(e15.0,clName = "CL_original.clusters")[[1]] == "511"]

cells <-names(cells)

cl511e15.0 <-  automaticCOTANObjectCreation(raw = getRawData(e15.0)[,cells],
                                       GEO = "Loom",
                                       sequencingMethod = "10X",
                                       sampleCondition = "cl511e15.0",
                                       cores = 10,saveObj = T,
                                       outDir = "Data/MouseCortexFromLoom/NotCTSingleClusters/",
                                       calcCoex = T)

```


```{r}
cl511e15.0GDI <- calculateGDI(cl511e15.0)

subsetGDIcl511e15.0 <- cl511e15.0GDI[cl511e15.0GDI$sum.raw.norm > 5,]
top.GDI.genes <- rownames(subsetGDIcl511e15.0[order(subsetGDIcl511e15.0$GDI,decreasing = T),])[1:25]

GDIPlot(cl511e15.0,genes = list("topGDI"=top.GDI.genes),GDIIn = cl511e15.0GDI)
```

```{r, eval=FALSE, echo=TRUE}
list.cl.CT <- cellsUniformClustering(cl511e15.0, GDIThreshold = 1.43, cores = 10L,
                         saveObj = TRUE, outDir = outDir)

cl511e15.0 <- addClusterization(cl511e15.0, clName = "split",
                              clusters = list.cl.CT$clusters,
                              coexDF = list.cl.CT$coex, override = TRUE)

table(list.cl.CT$clusters)
```
```{r, eval=FALSE, echo=TRUE}
c(mergedClusters, mergedCoexDF) %<-%
  mergeUniformCellsClusters(cl511e15.0, clusters = list.cl.CT$clusters,
                            GDIThreshold = 1.43, batchSize = 4L, cores = 10L,
                            saveObj = TRUE, outDir = outDir)

cl511e15.0 <- addClusterization(cl511e15.0, clName = "merge",
                              clusters = mergedClusters,
                              coexDF = mergedCoexDF,
                              override = TRUE)

saveRDS(cl511e15.0,paste0(outDir,"cl511e15.0.cotan.RDS"))

table(str_split(names(mergedClusters),pattern = "_",simplify = T)[,1],mergedClusters)
```



```{r}
data <- getNormalizedData(cl511e15.0)
data <- data[!rowSums(as.matrix(data)) < 1,]
data <- log(data*10000+1)

row_stdev <- apply(data, 1, sd, na.rm=TRUE)
row_stdev <- row_stdev[order(row_stdev,decreasing = T)]

genes.to.keep <- c(names(row_stdev[1:100]),top.GDI.genes)

data.small <- data[rownames(data) %in% genes.to.keep,]

#data <- t(as.matrix(data))
data.small <- t(as.matrix(data.small))


COTAN_Cl.code <- getClusterizationData(cl511e15.0)[[1]]
#COTAN_Cl.code[COTAN_Cl.code == "A549"] <- 0
#COTAN_Cl.code[COTAN_Cl.code == "CCL-185-IG"] <- 1

COTAN_Cl.code <- as.numeric(COTAN_Cl.code)-1

data.small <- cbind(data.small,COTAN_Cl.code)
data.small <- as.data.frame(data.small)

# Split the data into training and test set
set.seed(123)
training.samples <- data.small[,"COTAN_Cl.code"] %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- data.small[training.samples, ]
test.data <- data.small[-training.samples, ]
head(train.data)
```


```{r}
# Fit the model
model <- glm( COTAN_Cl.code ~., data = train.data, family = binomial,control = list(maxit = 50))
# Summarize the model
summary(model)
# Make predictions
probabilities <- model %>% predict(test.data, type = "response")
predicted.classes <- ifelse(probabilities > 0.5, "1", "0")
# Model accuracy
accuracy <- mean(predicted.classes == test.data$COTAN_Cl.code)
result.df[nrow(result.df),"Accuracy"] <- accuracy

accuracy

```

So also in this case a logistic regression confirm the difference in cell population detected by COTAN.

```{r}
cl511e15.0ClsMarkers <- findClustersMarkers(cl511e15.0,n = 200,method = "BH")
cl511e15.0ClsMarkers
```

```{r, echo=FALSE, eval=FALSE}
write.csv(cl511e15.0ClsMarkers,paste0(outDir,"cl511e15.0/ClustersMarkers.csv"))
write.csv(getClusterizationData(cl511e15.0)[[1]],file = paste0(outDir,"cl511e15.0/CellClusters.csv"))
```

## Summary

```{r}
result.df
```



------------------------------------------------------------------------

```{r}
Sys.time()
```

```{r}
sessionInfo()
```
