---
  title: "Probability score computation"
---

```{r}
library(rlang)
library(testthat)
library(ggplot2)
library(tibble)
library(zeallot)
library(Seurat)
library(aricode)
library(funtimes)
library(monocle3)
library(cluster)
library(COTAN)

tunings = c('default', 'antibody', 'celltypist')
tools = c('monocle', 'scanpy', 'seurat', 'scvi-tools', 'COTAN')


averageClustersDistance <- function(probMatrixDF,
                                    clustersList, numDim) {
  
  squareDist <- function(subMatrix) {
    subMatrix <- scale(subMatrix, scale = FALSE, center = TRUE)
    return(sum(rowSums(subMatrix^2)))
}
  
  pca <- PCAtools::pca(mat = probMatrixDF, rank = numDim, transposed = TRUE)

  residual_variance = 100 - sum(pca[["variance"]])

  normMatrix <- scale(x = pca[["rotated"]], center = TRUE, scale = TRUE)
  rownames(normMatrix) <- rownames(probMatrixDF)

  final_dimensions = dim(normMatrix)[2]

  sumDist <- 0.0
  for (cl in clustersList) {
    sumDist <- sumDist + squareDist(normMatrix[cl, , drop = FALSE])
  }
  score = sumDist / ncol(normMatrix) / (nrow(normMatrix) - length(clustersList))
  return(list("residual_variance" = residual_variance, "final_dimensions" = final_dimensions, "score" = score))
}



```
## PBMC1

```{r}
dataset = 'PBMC1'
celltypistPath = paste('Data/', dataset, '/celltypist/Immune_All_Low_probability_matrix.csv', sep='')
outPath = paste('Data/', dataset, '/probabilistic_score.csv', sep='')
```


Load CellTypist clusterization

```{r}
probMatrixDF <- read.csv(celltypistPath, header = TRUE)
probMatrixDF <- column_to_rownames(probMatrixDF, var = "X")
rownames(probMatrixDF) <- gsub("[.]", ":", rownames(probMatrixDF))
rownames(probMatrixDF) <- gsub("X10X", "10X", rownames(probMatrixDF))
rownames(probMatrixDF) <- substr(rownames(probMatrixDF), 1, nchar(rownames(probMatrixDF)) - 2)
```

Estimate number of dimensions

```{r}
prc <- PCAtools::pca(mat = probMatrixDF, rank = 50, transposed = TRUE)
plot(prc$sdev)
```

```{r}
nDims <- 20
```

Calculate the projection to spherical distribution [Malahanobis distance]:
the sum of square distances from center of sub-cluster in pca-projected space



```{r}
scores = data.frame()#"tool" = NA, "dataset" = dataset, "tuning" = NA, "score" = NA, "residual_variance" = NA, "final_dimensions" = NA)
for (tool in tools) {
  for (tuning in tunings) {
    labelsPath = paste('Data/', dataset, '/', tool, '/', tuning, '/clustering_labels.csv', sep='')
    labels <- read.csv(labelsPath, header = TRUE)
    clustersList<- split(as.character(labels$cell), labels$cluster)
    score_data = averageClustersDistance(probMatrixDF, clustersList,numDim = nDims)
    scores = rbind(
      scores,
      data.frame(
        "tool" = tool,
        "dataset" = dataset,
        "tuning" = tuning,
        "score" = score_data$score,
        "residual_variance" = score_data$residual_variance,
        "final_dimensions" = score_data$final_dimensions
      )
    )
  }
}
write.csv(scores, outPath, row.names = FALSE)
scores
```

```{r}
set.seed(42)
shuffledCells <- rownames(probMatrixDF)
rndDist <- function(n) {
  clusters <- set_names(rep(1:n, ceiling(9000/n))[1:nrow(probMatrixDF)], shuffledCells)
  cluster_list <- split(names(clusters), clusters)
  d <- averageClustersDistance(probMatrixDF, cluster_list,nDims)
  return(d$score)
}
sizes <- c(1:15, seq(20, 50, by = 5), seq(60, 100, by = 10),
           seq(150, 1000, by = 50), seq(1500, 9000, by = 500))
sizesToDists <- vapply(sizes,rndDist, numeric(1)) 
names(sizesToDists) <- sizes
pdf(paste('Data/', dataset, '/probability_score_dim_dependency.pdf', sep=''))
plot(names(sizesToDists), sizesToDists)
dev.off()
plot(names(sizesToDists), sizesToDists)
```
## PBMC2

```{r}
dataset = 'PBMC2'
celltypistPath = paste('Data/', dataset, '/celltypist/Immune_All_Low_probability_matrix.csv', sep='')
outPath = paste('Data/', dataset, '/probabilistic_score.csv', sep='')
```


Load CellTypist clusterization

```{r}
probMatrixDF <- read.csv(celltypistPath, header = TRUE)
probMatrixDF <- column_to_rownames(probMatrixDF, var = "X")
rownames(probMatrixDF) <- gsub("[.]", ":", rownames(probMatrixDF))
rownames(probMatrixDF) <- gsub("X10X", "10X", rownames(probMatrixDF))
rownames(probMatrixDF) <- substr(rownames(probMatrixDF), 1, nchar(rownames(probMatrixDF)) - 2)
```

Estimate number of dimensions

```{r}
prc <- PCAtools::pca(mat = probMatrixDF, rank = 50, transposed = TRUE)
plot(prc$sdev)
```

```{r}
nDims <- 20
```

Calculate the projection to spherical distribution [Malahanobis distance]:
the sum of square distances from center of sub-cluster in pca-projected space


```{r}
scores = data.frame()#"tool" = NA, "dataset" = dataset, "tuning" = NA, "score" = NA, "residual_variance" = NA, "final_dimensions" = NA)
for (tool in tools) {
  for (tuning in tunings) {
    labelsPath = paste('Data/', dataset, '/', tool, '/', tuning, '/clustering_labels.csv', sep='')
    labels <- read.csv(labelsPath, header = TRUE)
    clustersList<- split(as.character(labels$cell), labels$cluster)
    score_data = averageClustersDistance(probMatrixDF, clustersList,numDim = nDims)
    scores = rbind(
      scores,
      data.frame(
        "tool" = tool,
        "dataset" = dataset,
        "tuning" = tuning,
        "score" = score_data$score,
        "residual_variance" = score_data$residual_variance,
        "final_dimensions" = score_data$final_dimensions
      )
    )
  }
}
write.csv(scores, outPath, row.names = FALSE)
scores
```

```{r}
set.seed(42)
shuffledCells <- rownames(probMatrixDF)
rndDist <- function(n) {
  clusters <- set_names(rep(1:n, ceiling(9000/n))[1:nrow(probMatrixDF)], shuffledCells)
  cluster_list <- split(names(clusters), clusters)
  d <- averageClustersDistance(probMatrixDF, cluster_list,nDims)
  return(d$score)
}
sizes <- c(1:15, seq(20, 50, by = 5), seq(60, 100, by = 10),
           seq(150, 1000, by = 50), seq(1500, 9000, by = 500))
sizesToDists <- vapply(sizes, rndDist, numeric(1)) 
names(sizesToDists) <- sizes
pdf(paste('Data/', dataset, '/probability_score_dim_dependency.pdf', sep=''))
plot(names(sizesToDists), sizesToDists)
dev.off()
plot(names(sizesToDists), sizesToDists)
```


## PBMC3

```{r}
dataset = 'PBMC3'
celltypistPath = paste('Data/', dataset, '/celltypist/Immune_All_Low_probability_matrix.csv', sep='')
outPath = paste('Data/', dataset, '/probabilistic_score.csv', sep='')
```


Load CellTypist clusterization

```{r}
probMatrixDF <- read.csv(celltypistPath, header = TRUE)
probMatrixDF <- column_to_rownames(probMatrixDF, var = "X")
rownames(probMatrixDF) <- gsub("[.]", ":", rownames(probMatrixDF))
rownames(probMatrixDF) <- gsub("X10X", "10X", rownames(probMatrixDF))
rownames(probMatrixDF) <- substr(rownames(probMatrixDF), 1, nchar(rownames(probMatrixDF)) - 2)
```

Estimate number of dimensions

```{r}
prc <- PCAtools::pca(mat = probMatrixDF, rank = 50, transposed = TRUE)
plot(prc$sdev)
```

```{r}
nDims <- 20
```

Calculate the projection to spherical distribution [Malahanobis distance]:
the sum of square distances from center of sub-cluster in pca-projected space


```{r}
scores = data.frame()#"tool" = NA, "dataset" = dataset, "tuning" = NA, "score" = NA, "residual_variance" = NA, "final_dimensions" = NA)
for (tool in tools) {
  for (tuning in tunings) {
    labelsPath = paste('Data/', dataset, '/', tool, '/', tuning, '/clustering_labels.csv', sep='')
    labels <- read.csv(labelsPath, header = TRUE)
    clustersList<- split(as.character(labels$cell), labels$cluster)
    score_data = averageClustersDistance(probMatrixDF, clustersList,numDim = nDims)
    scores = rbind(
      scores,
      data.frame(
        "tool" = tool,
        "dataset" = dataset,
        "tuning" = tuning,
        "score" = score_data$score,
        "residual_variance" = score_data$residual_variance,
        "final_dimensions" = score_data$final_dimensions
      )
    )
  }
}
write.csv(scores, outPath, row.names = FALSE)
scores
```

```{r}
set.seed(42)
shuffledCells <- rownames(probMatrixDF)
rndDist <- function(n) {
  clusters <- set_names(rep(1:n, ceiling(9000/n))[1:nrow(probMatrixDF)], shuffledCells)
  cluster_list <- split(names(clusters), clusters)
  d <- averageClustersDistance(probMatrixDF, cluster_list,nDims)
  return(d$score)
}
sizes <- c(1:15, seq(20, 50, by = 5), seq(60, 100, by = 10),
           seq(150, 1000, by = 50), seq(1500, 9000, by = 500))
sizesToDists <- vapply(sizes, rndDist, numeric(1)) 
names(sizesToDists) <- sizes
pdf(paste('Data/', dataset, '/probability_score_dim_dependency.pdf', sep=''))
plot(names(sizesToDists), sizesToDists)
dev.off()
plot(names(sizesToDists), sizesToDists)
```

## PBMC4

```{r}
dataset = 'PBMC4'
celltypistPath = paste('Data/', dataset, '/celltypist/Immune_All_Low_probability_matrix.csv', sep='')
outPath = paste('Data/', dataset, '/probabilistic_score.csv', sep='')
```


Load CellTypist clusterization

```{r}
probMatrixDF <- read.csv(celltypistPath, header = TRUE)
probMatrixDF <- column_to_rownames(probMatrixDF, var = "X")
rownames(probMatrixDF) <- gsub("[.]", ":", rownames(probMatrixDF))
rownames(probMatrixDF) <- gsub("X10X", "10X", rownames(probMatrixDF))
rownames(probMatrixDF) <- substr(rownames(probMatrixDF), 1, nchar(rownames(probMatrixDF)) - 2)
```

Estimate number of dimensions

```{r}
prc <- PCAtools::pca(mat = probMatrixDF, rank = 50, transposed = TRUE)
plot(prc$sdev)
```

```{r}
nDims <- 20
```

Calculate the projection to spherical distribution [Malahanobis distance]:
the sum of square distances from center of sub-cluster in pca-projected space


```{r}
scores = data.frame()#"tool" = NA, "dataset" = dataset, "tuning" = NA, "score" = NA, "residual_variance" = NA, "final_dimensions" = NA)
for (tool in tools) {
  for (tuning in tunings) {
    labelsPath = paste('Data/', dataset, '/', tool, '/', tuning, '/clustering_labels.csv', sep='')
    labels <- read.csv(labelsPath, header = TRUE)
    clustersList<- split(as.character(labels$cell), labels$cluster)
    score_data = averageClustersDistance(probMatrixDF, clustersList,numDim = nDims)
    scores = rbind(
      scores,
      data.frame(
        "tool" = tool,
        "dataset" = dataset,
        "tuning" = tuning,
        "score" = score_data$score,
        "residual_variance" = score_data$residual_variance,
        "final_dimensions" = score_data$final_dimensions
      )
    )
  }
}
write.csv(scores, outPath, row.names = FALSE)
scores
```

```{r}
set.seed(42)
shuffledCells <- rownames(probMatrixDF)
rndDist <- function(n) {
  clusters <- set_names(rep(1:n, ceiling(9000/n))[1:nrow(probMatrixDF)], shuffledCells)
  cluster_list <- split(names(clusters), clusters)
  d <- averageClustersDistance(probMatrixDF, cluster_list,nDims)
  return(d$score)
}
sizes <- c(1:15, seq(20, 50, by = 5), seq(60, 100, by = 10),
           seq(150, 1000, by = 50), seq(1500, 9000, by = 500))
sizesToDists <- vapply(sizes, rndDist, numeric(1)) 
names(sizesToDists) <- sizes
pdf(paste('Data/', dataset, '/probability_score_dim_dependency.pdf', sep=''))
plot(names(sizesToDists), sizesToDists)
dev.off()
plot(names(sizesToDists), sizesToDists)
```


------------------------------------------------------------------------

```{r, eval=TRUE}
Sys.time()
```

```{r, eval=TRUE}
sessionInfo()
```