
---
  title: "Compare the various clusterizations of Forebrain Dorsal E13.5"
---

```{r}
#library(SingleCellExperiment)
#library(DuoClustering2018)
#library(tidyr)
library(rlang)
library(testthat)
library(ggplot2)
library(ggsankey) # remotes::install_github("davidsjoberg/ggsankey")
library(tibble)
library(zeallot)
library(Seurat)
library(aricode)
library(funtimes)
library(monocle3)
#library(COTAN)
devtools::load_all("~/dev/COTAN/COTAN/")

options(parallelly.fork.enable = TRUE)

setLoggingLevel(2)
setLoggingFile(file.path(".", "CompareClusterizations_ForebrainDorsal_E135.log"))
```
This function creates a sankey plot, given a COTAN object and the names
of 2 (or 3) clusterizations stored in the cells' metadata

```{r}
createSankey <- function(objCOTAN, clName1, clName2, clName3 = NULL,
                         space = 2, size = 4) {
  useThirdCluster <- !is.null(clName3)

  cl1 <- getClusters(objCOTAN, clName = clName1)
  if (!is.character(levels(cl1))) {
    cl1 <- niceFactorLevels(cl1)
  }

  cl2 <- getClusters(objCOTAN, clName = clName2)
  if (!is.character(levels(cl2))) {
    cl2 <- niceFactorLevels(cl2)
  }

  cl3 <- NULL
  if (useThirdCluster) {
    cl3 <- getClusters(objCOTAN, clName = clName3)
    if (!is.character(levels(cl3))) {
      cl3 <- niceFactorLevels(cl3)
    }
  }

  logThis(paste("Clusterization: ", clName1))
  logThis(paste("Number of clusters: ",
                nlevels(as.factor(cl1))), logLevel = 1L)
  logThis(paste("Top 10 clusters: ",
                paste0(head(sort(table(cl1), decreasing = TRUE), 10L),
                         collapse = ", ")), logLevel = 1L)

  logThis(paste("Clusterization: ", clName2))
  logThis(paste("Number of clusters: ",
                nlevels(as.factor(cl2))), logLevel = 1L)
  logThis(paste("Top 10 clusters: ",
                paste0(head(sort(table(cl2), decreasing = TRUE), 10L),
                         collapse = ", ")), logLevel = 1L)

  if (useThirdCluster) {
    logThis(paste("Clusterization: ", clName3))
    logThis(paste("Number of clusters: ",
                  nlevels(as.factor(cl3))), logLevel = 1L)
    logThis(paste("Top 10 clusters: ",
                  paste0(head(sort(table(cl3), decreasing = TRUE), 10L),
                           collapse = ", ")), logLevel = 1L)
  }

  df <- data.frame(cbind(factorToVector(cl1), factorToVector(cl2)))
  colnames(df) <- paste0(c(clName1, clName2), ".clusters")
  if (useThirdCluster) {
    df <- cbind(df, factorToVector(cl3))
    colnames(df)[[3]] <- paste0(clName3, ".clusters")
  }

  crossTable <- list(table(df[, c(1,2)]))
  if (useThirdCluster) {
    crossTable <- append(crossTable, list(table(df[, c(1,3)]), table(df[, c(2,3)])))
  }

  table2 <- df %>% make_long(1:ncol(df))

  sankeyPlot <-
    ggplot(table2,
           aes(x = x,
               next_x = next_x,
               node = node,
               next_node = next_node,
               fill = factor(node),
               label = node)) +
    geom_alluvial(flow.alpha = 0.75, node.color = 0, space = space) +
    geom_alluvial_text(space = space) +
#    geom_alluvial_label(space = space) +
    scale_fill_viridis_d(option = "H", alpha = 0.95, begin = 0.2, end = 0.8) +
    theme_alluvial(base_size = 16, base_line_size = size) +
    theme(legend.position = "none")
#    theme(plot.margin = unit(c(0,0,0,0), "mm"))
#    theme(legend.box.spacing = unit(0, "mm"))
#    labs(title = "Alluvial diagram using ggplot")
#    labs(subtitle = "using  David Sjoberg's ggsankey package")
#    labs(caption = "@techanswers88")

  return(list("tables" = crossTable, "plot" = sankeyPlot))
}
```


Loading the COTAN object

```{r}
fb135Obj <- readRDS(file = file.path(".", "e13.5_ForebrainDorsal.cotan.RDS"))

sampleCondition <- getMetadataElement(fb135Obj, datasetTags()[["cond"]])

logThis(sampleCondition)

allClust <- getClusterizations(fb135Obj)

logThis("")
logThis(paste("Number of cells:", getNumCells(fb135Obj)))

logThis("")
logThis("Available clusterizations:")
logThis(paste0(allClust, collapse = ", "))

clSizes <- vapply(allClust, function(x) { nlevels(getClusters(fb135Obj, x))}, integer(1L))
clSizes
logThis(paste(names(clSizes), clSizes, sep = ": ", collapse = ", "))
```

Load CellTypist clusterization

```{r}
probMatrixDF <- read.csv(file.path(".", "Labeling", "E135_Devel_Mouse_Brain_probability_matrix.csv"), header = TRUE)
probMatrixDF <- column_to_rownames(probMatrixDF, var = "X")
rownames(probMatrixDF) <- gsub("[.]", ":", rownames(probMatrixDF))
rownames(probMatrixDF) <- gsub("X10X", "10X", rownames(probMatrixDF))

cells_to_keep <- rownames(probMatrixDF)[rownames(probMatrixDF) %in% getCells(fb135Obj)]
assert_that(identical(cells_to_keep, getCells(fb135Obj)))
```

Estimate number of dimensions

```{r}
prc <- irlba::prcomp_irlba(probMatrixDF, n = 50)

nDims <- 20
plot(prc$sdev[10:nDims])
prc$totalvar - sum(prc$sdev[1:nDims]^2)
```

Calculate the projection to spherical distribution [Malahanobis distance]

```{r}
pca <- irlba::prcomp_irlba(probMatrixDF, n = nDims)

logThis(paste("Residual variance: ", prc$totalvar - sum(prc$sdev[1:nDims]^2)))

normMatrix <- scale(x = pca$x, center = TRUE, scale = TRUE)
rownames(normMatrix) <- rownames(probMatrixDF)

logThis(paste("Final dimensions:", paste0(dim(normMatrix), collapse = ", ")))
```

The sum of square distances from center of sub-cluster in pca-projected space

```{r}
squareDistanceFromCellTypist <- function(subMatrix) {
  subMatrix <- scale(subMatrix, scale = FALSE, center = TRUE)
  return(sum(rowSums(subMatrix^2)))
}
```


```{r}
averageClustersDistance <- function(normMatrix, clusterization) {
  clustersList <- toClustersList(clusterization)
  sumDist <- 0.0
  for (cl in clustersList) {
    sumDist <- sumDist + squareDistanceFromCellTypist(normMatrix[cl, , drop = FALSE])
  }
  return(sumDist / ncol(normMatrix) / (nrow(normMatrix) - length(clustersList)))
}
```


```{r eval=FALSE}
set.seed(155)
shuffledCells <- sample(getCells(fb135Obj))

rndDist <- function(n) {
  clusters <- set_names(rep(1:n, ceiling(3000/n))[1:getNumCells(fb135Obj)], shuffledCells)
  d <- averageClustersDistance(normMatrix, clusters)
  return(d)
}

sizes <- c(1:15, seq(20, 50, by = 5), seq(60, 100, by = 10),
           seq(150, 1000, by = 50), seq(1500, 5000, by = 500))

sizesToDists <- vapply(sizes, rndDist, numeric(1))
names(sizesToDists) <- sizes

sizesToDists

plot(names(sizesToDists), sizesToDists)
```



```{r}
avgDists <- vapply(allClust, function(clName) {
  averageClustersDistance(normMatrix, getClusters(fb135Obj, clName))
}, numeric(1))

avgDists
```



```{r}
tableAndPlot <- createSankey(fb135Obj, "cell.typist", "merge", space = 2, size = 2)

tableAndPlot
```

```{r}
allMutualScores <- function(objCOTAN, clName) {
  cbind(
    "NMI" = vapply(allClust, function(clName2) {
        NMI(getClusters(objCOTAN, clName = clName),
            getClusters(objCOTAN, clName = clName2),
            variant = "sum")
      }, numeric(1)),
  
    "AMI" = vapply(allClust, function(clName2) {
        AMI(getClusters(objCOTAN, clName = clName),
            getClusters(objCOTAN, clName = clName2))
      }, numeric(1)),

    "ARI" = vapply(allClust, function(clName2) {
        ARI(getClusters(objCOTAN, clName = clName),
            getClusters(objCOTAN, clName = clName2))
      }, numeric(1)),

    "purity" = vapply(allClust, function(clName2) {
        purity(getClusters(objCOTAN, clName = clName),
               getClusters(objCOTAN, clName = clName2))[["pur"]]
      }, numeric(1))
  )
}
```


```{r}
cellTypistScores <- allMutualScores(fb135Obj, clName = "cell.typist")

cellTypistScores
```

```{r}
originalSubclassScores <- allMutualScores(fb135Obj, clName = "original.subclass")

originalSubclassScores
```


```{r}
originalClusterScores <- allMutualScores(fb135Obj, clName = "original.clusters")

originalClusterScores
```


```{r}

comparativeScores <- cbind(clSizes, avgDists,
                           originalClusterScores[, "AMI"], originalClusterScores[ , "ARI"],
                           cellTypistScores[ , "AMI"], cellTypistScores[ , "ARI"])

colnames(comparativeScores) <- c("NumClusters", "NPD",
                                 "AMI_OriginalClusters", "ARI_OriginalClusters",
                                 "AMI_CellTypist", "ARI_CellTypist")

comparativeScores

write.csv(comparativeScores, file = "./ForebrainDorsal_E135_ClustersComparativeScores.csv")
```
