---
title: "GDI_Increment_From_Mixing"
author: "Marco Fantozzi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preamble

```{r preamble}
library(assertthat)
library(rlang)
library(scales)
library(ggplot2)
library(zeallot)
library(data.table)
#library(COTAN)

#options(parallelly.fork.enable = TRUE)

inDir <- file.path(".")

#setLoggingLevel(2)
#setLoggingFile(file.path(inDir, "MixingClustersGDI_ForebrainDorsal.log"))

outDir <- file.path(inDir, "GDI_Sensitivity")
if (!file.exists(outDir)) {
  dir.create(outDir)
}
```

```{r eval=TRUE, echo=FALSE}
selectedClusters <- list(
    "E13.5" = c("432", "187", "434", "184", "437", "510"),
    "E15.0" = c("432", "509", "510", "508", "428", "434", "437"),
    "E17.5" = c("516", "505")
)
```

## Load calculated data for analysis

```{r echo=FALSE, eval=TRUE}
resMix05 <- readRDS(file.path(outDir, paste0("GDI_with_05%_Mixing.RDS")))
resMix10 <- readRDS(file.path(outDir, paste0("GDI_with_10%_Mixing.RDS")))
resMix20 <- readRDS(file.path(outDir, paste0("GDI_with_20%_Mixing.RDS")))
resMix40 <- readRDS(file.path(outDir, paste0("GDI_with_40%_Mixing.RDS")))
resMix80 <- readRDS(file.path(outDir, paste0("GDI_with_80%_Mixing.RDS")))

# create an aligned baseline
resMix00 <- resMix05
resMix00[["MixingFraction"]] <- 0.0
resMix00[["GDI"]] <- resMix00[["GDI"]] - resMix00[["GDIIncrement"]]
resMix00[["GDIIncrement"]] <- 0.0
```


## Recall cluster distance and add it to the results

```{r}
zeroOneAvg <- readRDS(file.path(inDir, "ClustersDistances", "allZeroOne.RDS"))
distZeroOne <- as.matrix(parDist(t(zeroOneAvg), method = "hellinger", diag = TRUE, upper = TRUE))^2

distZeroOneLong <- rownames_to_column(as.data.frame(distZeroOne), var = "MainCluster")
distZeroOneLong <-pivot_longer(distZeroOneLong,
                               cols = !MainCluster,
                               names_to = "OtherCluster", 
                               values_to = "Distance")

distZeroOneLong <- as.data.frame(distZeroOneLong[distZeroOneLong[["Distance"]] != 0.0, ])

assert_that(identical(distZeroOneLong[, 1:2], resMix20[, 1:2]))

perm <- order(distZeroOneLong[["Distance"]])
```

```{r}
# Scatter plot of the effective increment at 40% mixing [Y]
# against estimated distance [X]
distDF <- cbind(distZeroOneLong[, "Distance", drop = FALSE],
                sqrt(distZeroOneLong[, "Distance", drop = FALSE]))
colnames(distDF) <- c("Distance", "DistanceSqrt")

D2IPlot <- ggplot(cbind(resMix40, distDF),
                  aes(x=Distance, y=GDIIncrement)) +
             geom_point() +
             geom_smooth(method=lm, formula = y ~ x) 
           # +  xlim(0, 1.5) + ylim(0, 1.5) + coord_fixed()

plot(D2IPlot)
```


## Merge all data and plot it using the distance as discriminant

```{r}
allRes <- rbind(resMix05[perm, ], resMix10[perm, ], resMix20[perm, ], resMix40[perm, ], resMix80[perm, ])
allRes <- cbind(allRes, "Distance" = rep(distZeroOneLong[["Distance"]][perm], 5))
rownames(allRes) <- NULL
allRes <- cbind(allRes, "ClusterPair" = rep.int(c(1:210),5))


allResWithBase <- cbind(resMix00[perm, ], "Distance" = distZeroOneLong[["Distance"]][perm])
rownames(allResWithBase) <- NULL
allResWithBase <- cbind(allResWithBase, "ClusterPair" = c(1:210))
allResWithBase <- rbind(allResWithBase, allRes)

assert_that(identical(allRes[, 4], allResWithBase[211:1260, 4]))
```

```{r}
IScPlot <- ggplot(allRes, aes(x=MixingFraction, y=GDIIncrement, color=Distance,)) +
  geom_point() +
  scale_color_continuous(type = "viridis") +
  scale_x_log10()

plot(IScPlot)

GScPlot <- ggplot(allResWithBase, aes(x=MixingFraction, y=GDI, color=Distance,)) +
  geom_point() +
  scale_color_continuous(type = "viridis")

plot(GScPlot)
```


```{r}
reOrder <- function(df, numBlocks) {
  blockLength <- nrow(df) / numBlocks
  permut <- rep(1:blockLength, each = numBlocks) +
              rep(seq(0, nrow(df) - 1, by = blockLength), times = numBlocks)
  return(df[permut, ])
}

allRes2 <- reOrder(allRes, 5)
allResWithBase2 <- reOrder(allResWithBase, 6)
```

```{r}
rng <- c(1,210)
#rng <- c(1,42)
#rng <- c(43,84)
#rng <- c(85,126)
#rng <- c(127,168)
#rng <- c(169,210)

ILinesPLot <- ggplot(allRes2[allRes2[["ClusterPair"]] %between% rng, ],
                     aes(x = MixingFraction, y = GDIIncrement,
                         color = (ClusterPair - 1) %/% 42 + 0.5)) + 
  geom_path(aes(group = ClusterPair)) +
  theme(legend.position = "none") +
  #scale_x_log10() + 
  scale_colour_stepsn(colours = hcl.colors(5, palette = "Dark 2")[5:1])

plot(ILinesPLot)


GLinesPLot <- ggplot(allResWithBase2[allResWithBase2[["ClusterPair"]] %between% rng, ],
                     aes(x = MixingFraction, y = GDI,
                         color = (ClusterPair - 1) %/% 42 + 0.5)) + 
  geom_path(aes(group = ClusterPair)) +
  theme(legend.position = "none") +
  #scale_x_log10() + 
  scale_colour_stepsn(colours = hcl.colors(5, palette = "Dark 2")[5:1]) 
#  geom_line(data = data.frame(cbind(MixingFraction = c(0,0.8), GDI = c(1.4,1.4))),
#            aes(x = MixingFraction, y = GDI))


plot(GLinesPLot)
```

```{r}
mg <- function(mixings) {
  res <- mixings
  res[res !=0 ] <- ceiling(log2(round(res[res !=0 ] * 40)))
  return(res)
}

rng <- c(1,42)
#rng <- c(43,84)
#rng <- c(85,126)
#rng <- c(127,168)
#rng <- c(169,210)

IBoxPlot <- ggplot(allRes[allRes[["ClusterPair"]] %between% rng, ],
                   aes(x=mg(MixingFraction), y=GDIIncrement, color=Distance,
                              group = mg(MixingFraction))) +
  geom_boxplot()

plot(IBoxPlot)

GBoxPlot <- ggplot(allResWithBase[allResWithBase[["ClusterPair"]] %between% rng, ],
                   aes(x=mg(MixingFraction), y=GDI, color=Distance,
                              group = mg(MixingFraction))) +
  geom_boxplot()

plot(GBoxPlot)
```



## Load calculated data about multi clusters cases for analysis

```{r eval=TRUE, echo=FALSE}
resMix20_2 <- readRDS(file.path(outDir, paste0("GDI_with_20%_Mixing_from_Pairs.RDS")))
```

## Compare estimated vs real GDI increment

```{r}
# Scatter plot of the effective increment [Y] against estimated increment [X]
pg <- ggplot(resMix20_2, aes(x=PredictedGDIIncrement, y=GDIIncrement)) +
  geom_point() +
  geom_smooth(method=lm, formula = y ~ x + 0) +
  coord_fixed() +
  xlim(0, 1.5) + ylim(0, 1.5)
  #scale_x_log10() + scale_y_log10() 

plot(pg)
```

The plot shows that having the 20% extraneous cells in the mixture coming from 
multiple clusters does not affect significantly the sensitivity of the GDI to
score cluster uniformity.
