---
title: "Original Clusters Distance Evaluation"
author: "seriph78"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(COTAN)
library(stringr)
library(tidyverse)
library(ComplexHeatmap)

options(parallelly.fork.enable = TRUE)

inDir <- file.path(".")

setLoggingLevel(2)
setLoggingFile(file.path(inDir, "ClustersDistance_ForebrainDorsal.log"))

outDir <- file.path(inDir, "ClustersDistances")
if (!file.exists(outDir)) {
  dir.create(outDir)
}
```

## Loading all COTAN Objects

This is done on the La Manno Mouse Brain dataset (2021)

```{r load all}
fb135Obj <- readRDS(file = file.path(inDir, "e13.5_ForebrainDorsal.cotan.RDS"))
getMetadataElement(fb135Obj, datasetTags()[["cond"]])

fb150Obj <- readRDS(file = file.path(inDir, "e15.0_ForebrainDorsal.cotan.RDS"))
getMetadataElement(fb150Obj, datasetTags()[["cond"]])

fb175Obj <- readRDS(file = file.path(inDir, "e17.5_ForebrainDorsal.cotan.RDS"))
getMetadataElement(fb175Obj, datasetTags()[["cond"]])
```

## Relevant clusters lists

```{r relevant clusters, echo=FALSE}
selected <- list(
    "E13.5" = c("432", "187", "434", "184", "437", "510"),
    "E15.0" = c("432", "509", "510", "508", "428", "434", "437"),
    "E17.5" = c("516", "505")
)

objSelector <- function(dsName) {
  if (dsName == "E13.5") {
    return(fb135Obj)
  } else if (dsName == "E15.0") {
    return(fb150Obj)
  } else if (dsName == "E17.5") {
    return(fb175Obj)
  } else {
   stop("Wrong dataset name") 
  }
}
```

## Define function to *merge* clusters from different datasets

```{r}
allGenes <- union(union(getGenes(fb135Obj), getGenes(fb150Obj)), getGenes(fb175Obj))

# This function completes the raw data so that the genes list
# matches the `allGenes` above
fillGenes <- function(m, prefixName) {
  # find the missing genes in the given matrix m
  missingGenes <- !(allGenes %in% rownames(m))
  # create a matrix with the right sizes for the missing genes
  # and set their values all to zero
  fillM <- matrix(0, nrow = sum(missingGenes), ncol = ncol(m))
  rownames(fillM) <- allGenes[missingGenes]
  # append the new matrix to the original and reorder the rows
  # to match overall genes order
  retM <- rbind(m, fillM)
  retM <- retM[allGenes, ]
  # assign unique identifiers to the columns, so that one can discriminate
  # the source even after merging with other matrices
  colnames(retM) <- paste0(prefixName, "_", colnames(retM))
  return(retM)
}
```

## Extract the raw data for each cluster
### Made a list after gene-completing each of them

```{r}
allClustersRawData <- list()
for (dsName in names(selected)) {
  clList <- toClustersList(getClusters(objSelector(dsName),
                                       clName = "original.clusters"))
  for (clName in selected[[dsName]]) {
    fullName <- paste0(dsName, ":", clName)
    fullRawData <- fillGenes(getRawData(objSelector(dsName))[, clList[[clName]]],
                             fullName)
    allClustersRawData <- append(allClustersRawData, list(fullRawData))
    names(allClustersRawData)[length(allClustersRawData)] <- fullName
  }
}

assert_that(identical(rownames(allClustersRawData[[1]]), allGenes))

# delete the COTAN objects
rm(fb135Obj, fb150Obj, fb175Obj)
```


## Defining the two distances

To roughly determine the cluster distances we decided to test two simple Euclidean distances:

1.  over the mean of the 0/1 raw matrix

2.  over the $1-e^{-\lambda}$ where $\lambda$ is the average expression for the genes.

```{r, echo=TRUE, eval=FALSE}
allZeroOne <- data.frame(row.names = allGenes)
allProbOne <- data.frame(row.names = allGenes)

for (clName in names(allClustersRawData)) {
  print("Calculating statistics for custer:", clName)
  
#  cl <- str_split(Fl,pattern = "_",simplify = T)[1]
  obj <- COTAN(allClustersRawData[[clName]])

  # for each gene get average presence in a cell 
  ZeroOne <- rowMeans(getZeroOneProj(obj))
  allZeroOne <- cbind(allZeroOne, ZeroOne)
  colnames(allZeroOne)[ncol(allZeroOne)] <- clName

  # for each gene get the average probability of presence in a cell
  obj <- estimateLambdaLinear(obj)
  ProbOne <- 1 - exp(-getLambda(obj)) 
  allProbOne <- cbind(allProbOne, ProbOne)
  colnames(allProbOne)[ncol(allProbOne)] <- clName
}

head(allZeroOne)
head(allProbOne)
```

```{r, eval=FALSE, echo=TRUE}
saveRDS(allZeroOne, file.path(outDir, "allZeroOne.RDS"))
saveRDS(allProbOne, file.path(outDir, "allProbOne.RDS"))
```

```{r, echo=FALSE, eval=TRUE}
allZeroOne <- readRDS(file.path(outDir, "allZeroOne.RDS"))
allProbOne <- readRDS(file.path(outDir, "allProbOne.RDS"))
```

```{r}
distanceProbOne <- as.matrix(dist(t(allProbOne), diag = TRUE, upper = TRUE))

Heatmap(distanceProbOne,
        name = "Lambda\ndistance", 
        cell_fun = function(j, i, x, y, width, height, fill) 
        {
          grid.text(sprintf("%.1f", distanceProbOne[i, j]), x, y, gp = gpar(fontsize = 10))
        },
        show_row_dend = FALSE,
        show_row_names = FALSE)
```

```{r}

distanceZeroOne <- as.matrix(dist(t(allZeroOne), diag = TRUE, upper = TRUE))

Heatmap(distanceZeroOne,
        name = "ZeroOne\ndistance", 
        cell_fun = function(j, i, x, y, width, height, fill) 
        {
          grid.text(sprintf("%.1f", distanceZeroOne[i, j]), x, y, gp = gpar(fontsize = 10))
        },
        show_row_dend = F, 
        show_row_names = F)
```

The distances are very similar with the Zero One a little lower as values...

```{r}
distanceProbOneLong <- rownames_to_column(as.data.frame(distanceProbOne),
                                          var = "Cl.1")
distanceProbOneLong <- pivot_longer(distanceProbOneLong,
                                    cols = !Cl.1,
                                    names_to = "Cl.2", 
                                    values_to = "Lambda.Dist")

distanceZeroOneLong <- rownames_to_column(as.data.frame(distanceZeroOne),
                                          var = "Cl.1")
distanceZeroOneLong <-pivot_longer(distanceZeroOneLong,
                                   cols = !Cl.1,
                                   names_to = "Cl.2", 
                                   values_to = "ZeroOne.Dist")

distanceAll <- merge(distanceZeroOneLong, distanceProbOneLong,
                     by = c("Cl.1", "Cl.2"), all = TRUE)

ggplot(as.data.frame(distanceAll), aes(x = ZeroOne.Dist, y = Lambda.Dist)) + geom_point()
```

So for what we are using it is the same.

The key information is that we can define couple of, in one case, triplets of very near clusters:

1.  Cl510e13.5 and Cl510e15.0
2.  Cl516e17.5 and Cl505e17.5
3.  Cl509e15.0 and Cl508e15.0 which are also similar to Cl432e13.5 and Cl432e15.0
4.  Cl432e13.5 and Cl432e15.0
5.  Cl437e13.5 and Cl437e15.0
6.  Cl428e15.0, Cl434e15.0 and Cl434e13.5

Based on these distances we can consider three thresholds (with number of cluster pair):

```{r}
quantile(distanceZeroOne[distanceZeroOne != 0.0], probs = seq(from = 0, to = 1, by = 0.05))
```


1.  less than 7 (first quintile)

```{r}
sum(distanceZeroOne < 7.17 & distanceZeroOne != 0.0) / sum(distanceZeroOne != 0.0)
```

2. between 9 and 11 (around median)

```{r}
sum(distanceZeroOne > 9.4 & distanceZeroOne < 11.9) / sum(distanceZeroOne != 0.0)
```

3. more than 14.7 (last quintile)

```{r}
sum(distanceZeroOne > 14.6) / sum(distanceZeroOne != 0.0)
```

------------------------------------------------------------------------

```{r}
sessionInfo()
```
